### 2021-05-11 paper/t1_appendix.tex
update ms

%% \begin{deluxetable}{} command tell LaTeX how many columns
%% there are and how to align them.
%\startlongtable
\begin{deluxetable*}{lll}

%% Keep a portrait orientation

%% Over-ride the default font size
%% Use Default (12pt)
\tabletypesize{\scriptsize}
%\tabletypesize{\small}

%% Use \tablewidth{?pt} to over-ride the default table width.
%% If you are unhappy with the default look at the end of the
%% *.log file to see what the default was set at before adjusting
%% this value.

%% This is the title of the table.
\tablecaption{Young, Age-dated, and Age-dateable Stars Within the
  Nearest Few Kiloparsecs ($\texttt{v0.5}$ of the CDIPS Target
  List).}
\label{tab:v05}

%% This command over-rides LaTeX's natural table count
%% and replaces it with this number.  LaTeX will increment 
%% all other tables after this table based on this number
%\tablenum{3}

%% The \tablehead gives provides the column headers.  It
%% is currently set up so that the column labels are on the
%% top line and the units surrounded by ()s are in the 
%% bottom line.  You may add more header information by writing
%% another line between these lines. For each column that requries
%% extra information be sure to include a \colhead{text} command
%% and remember to end any extra lines with \\ and include the 
%% correct number of &s.
\tablehead{
  \colhead{Parameter} &
  \colhead{Example Value} &
  \colhead{Description}
}

%% All data must appear between the \startdata and \enddata commands
%
% paste from
% /Users/luke/Dropbox/proj/rudolf/results/tables/v05_main_tableheader.tex
% via drivers/write_v05_main_tableheader.py
\startdata
         \texttt{source\_id} &                                         1709456705329541504 &                                             Gaia DR2 source identifier. \\
                 \texttt{ra} &                                                     247.826 &                                         Gaia DR2 right ascension [deg]. \\
                \texttt{dec} &                                                      79.789 &                                             Gaia DR2 declination [deg]. \\
           \texttt{parallax} &                                                      35.345 &                                                Gaia DR2 parallax [mas]. \\
    \texttt{parallax\_error} &                                                       0.028 &                                    Gaia DR2 parallax uncertainty [mas]. \\
               \texttt{pmra} &                                                      94.884 &     Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr${^-1}$]. \\
              \texttt{pmdec} &                                                     -86.971 &                 Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr${^-1}$]. \\
 \texttt{phot\_g\_mean\_mag} &                                                        6.85 &                                                 Gaia DR2 $G$ magnitude. \\
\texttt{phot\_bp\_mean\_mag} &                                                       6.409 &                                     Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
\texttt{phot\_rp\_mean\_mag} &                                                       7.189 &                                     Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
            \texttt{cluster} &                 Uma,IR\_excess,NASAExoArchive\_ps\_20210506 &                                  Comma-separated cluster or group name. \\
                \texttt{age} &                                                nan,nan,9.48 & Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years. \\
          \texttt{mean\_age} &                                                        9.48 &                          Mean (ignoring NaNs) of $\texttt{age}$ column. \\
      \texttt{reference\_id} &      Ujjwal2020,CottenSong2016,NASAExoArchive\_ps\_20210506 &                          Comma-separted provenance of group membership. \\
 \texttt{reference\_bibcode} & 2020AJ....159..166U,2016ApJS..225...15C,2013PASP..125..989A &                  ADS bibcode corresponding to $\texttt{reference\_id}$. \\
\enddata

%% Include any \tablenotetext{key}{text}, \tablerefs{ref list},
%% or \tablecomments{text} between the \enddata and 
%% \end{deluxetable} commands

%% General table comment marker
\tablecomments{
Table~\ref{tab:v05} is
published in its entirety in a machine-readable format.   This table is a
concatenation of the studies listed in Table~\ref{tab:metadata}.
One entry is
shown for guidance regarding form and content. 
In this particular example, the star has a
cold Jupiter on a 16 year orbit, HD 150706b \citep{2012AA...545A..55B}.
An infrared excess has been reported \citep{CottenSong2016}, and the star was identified by \citet{Ujjwal2020} as a candidate UMa moving group
member ($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).
The star's RV activity and TESS rotation period corroborate its youth.
}
\vspace{-0.5cm}
\end{deluxetable*}