### 2021-06-21 paper/authors.tex
add ttv vs local slope analysis

\correspondingauthor{L.\,G.\,Bouma}
\email{bouma.luke@gmail.com}

\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\author[0000-0002-2792-134X]{J. L. Curtis}
\affiliation{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
\affiliation{Department of Astrophysics, American Museum of Natural History, Central Park West, New York, NY 10024, USA}

\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{Department of Earth and Space Science, Osaka University, Osaka 560-0043, Japan}