### 2021-10-08 paper/ms.tex
update ms

\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

\newcommand{\cn}{$\delta$\ Lyr\ cluster} % cluster name
\newcommand{\sn}{Kepler\,1627} % star system name (binary)
\newcommand{\pn}{Kepler\,1627Ab} % planet name

\newcommand{\clusterage}{$38^{+6}_{-5}$\,Myr} %

% gaia target sample numbers
\newcommand{\nkinematic}{1{,}201} % deltalyr_kc19_cleansubset_withreddening.csv
% all from oneoff_calculate_rotation_numbers.py
\newcommand{\noriginal}{3{,}071} % Vizier; KC19 Theia 73 member count
\newcommand{\nwithtess}{924} % How many stars have TESS data with G<17 and (Bp-Rp)0 > 0.5?
\newcommand{\nforgyro}{1842} % How many stars have G<17 and (Bp-Rp)0 > 0.5?
\newcommand{\nkinwithtess}{391} % How many kinematically selected stars have TESS data?
\newcommand{\nkinwithtessandcrowding}{192} % How many kinematically selected stars have TESS data and meeting the crowding cutoff (nequal<=0, nclose<=1)?
\newcommand{\nkindefaultcleaning}{145} % How many of these meet the "defaultcleaning" criteria (LSP>0.1, Prot<15, and crowding cut)?
\newcommand{\nrotgood}{135} % nkindefaultcleaning, minus outliers.
\newcommand{\nfracprot}{70} % fraction nrotgood/nkinwithtessandcrowding

%
% Symbols
%
\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\ms}{\,m\,s$^{-1}$}
\newcommand{\bpmrpo}{(G_{\rm BP}-G_{\rm RP})_0}
\newcommand{\bpmrp}{G_{\rm BP}-G_{\rm RP}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{AAS Journals}
\shorttitle{Kepler\,1627}

\begin{document}

\title{
  A 38 Million Year Old Mini-Neptune in the Kepler Field
}

%\suppressAffiliations
%\NewPageAfterKeywords
\input{authors.tex}

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.8\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler space telescope because it is nearby
  ($d=329\,{\rm pc}$) and it resembles the Sun.  Here we show using
  Gaia kinematics, TESS stellar rotation periods, and spectroscopic
  lithium abundances that Kepler\,1627 is a member of the \clusterage\
  old $\delta$~Lyr cluster.  To our knowledge, this makes
  Kepler\,1627Ab the youngest planet with a precise age yet found by
  the main Kepler mission.  The Kepler photometry shows two
  peculiarities: the average transit profile is asymmetric, and the
  individual transit times are correlated with the local light curve
  slope.  We discuss possible explanations for each anomaly.  More
  broadly, the $\delta$~Lyr cluster
  is one of $\sim$10$^3$ coeval groups whose properties have been clarified
  by Gaia.  Many other exoplanet hosts are potentially kinematic members
  of these clusters; their candidate memberships can be verified by
  measuring stellar rotation periods in the TESS full frame images.
\end{abstract}

\keywords{
  exoplanet evolution (491),
  open star clusters (1160),
	stellar ages (1581)
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

While thousands of exoplanets have been discovered orbiting nearby
stars, the vast majority of them are several billion years old.  This
makes it difficult to test origin theories for the different families
of planets, since many evolutionary processes are expected to operate
on timescales of less than 100 million years.

For instance, the ``mini-Neptunes'', thought to be made of metal
cores, silicate mantles \citep{kite_atmosphere_2020}, and extended
hydrogen-dominated atmospheres, are expected to shrink
in size by factors of several over their first $10^8$ years.
Specifically, in the models of \citet{owen_atmospheres_2016} and
\citet{owen_constraining_2020}, the $\approx5\,M_\oplus$ planets start
with sizes of 4--12\,$R_\oplus$ shortly after the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
is expected to occur within the first few million years after the disk
disperses \citep{ikoma_situ_2012}, stellar
irradiation and internal heat can also power gradual outflows
which, if strong enough, can deplete or entirely strip the atmospheric
envelope \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
Discovering young planets, measuring their masses, and detecting their
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of mature
exoplanets \citep{Fulton_et_al_2017}.

The K2 and TESS missions have now enabled the detection of about ten
close-in planets younger than 100 million years, all smaller than
Jupiter
\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020,martioli_aumicbc_2021}.
The Kepler mission however has not yielded any planets with precise
ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
during the main Kepler mission (2009--2013), only four open clusters
were known in the Kepler field,
%: NGC\,6866, NGC\,6811, NGC\,6819, and NGC\,6791,
with ages spanning 0.7\,Gyr to 9\,Gyr \citep{meibom_kepler_2011}.
Though isochronal, gyrochronal, and lithium-based analyses suggest
that younger Kepler planets do exist
\citep{berger_identifying_2018,david_sizes_2021}, accurate and precise
age measurements typically require an ensemble of stars.  Fortunately,
recent analyses of the Gaia data have greatly expanded our knowledge
of cluster memberships \citep[{\it
e.g.},][]{CantatGaudin2018a,Zari2018,KounkelCovey2019,Meingast2021,Kerr2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

Matching our young star list against stars observed by Kepler revealed
that Kepler observed a portion of the \cn\ (Stephenson-1; Theia~73).
More specifically, a clustering analysis of the Gaia data by
\citet{KounkelCovey2019} reported that Kepler~1627 (KIC 6184894; KOI
5245; TIC 120105470) is a \cn\ member.  Given the previous statistical
validation of the close-in mini-Neptune Kepler 1627b
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we begin by examining the properties of the cluster
(Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr cluster is
\clusterage\ old, and in Section~\ref{sec:stars} show that \sn\ is
both a binary and also a member of the cluster.  Focusing on the
planet (Section~\ref{sec:planet}), we confirm that despite the
existence of the previously unreported M2.5V companion, hereafter
Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A.
Based on a correlation between the planetary transit times and the
local light curve slope, the orbit may also be prograde.  We conclude
by highlighting broader implications for our ability to age-date a
larger sample of planets (Section~\ref{sec:conc}).

\section{The Cluster}
\label{sec:cluster}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Galactic positions and tangential velocities of stars in the
    $\delta$\,Lyr cluster.} Points are reported cluster members from
    \citet{KounkelCovey2019}.  The tangential velocities relative to
    \sn\ (lower-right) are computed assuming that every star has the
    same three-dimensional spatial velocity as \sn.  Our analysis
    considers stars (black points) in the spatial and kinematic
    vicinity of Kepler\,1627 (yellow star).  The question of whether
    the other candidate cluster members (gray points) are part of the
    cluster is outside our scope.  The location of the Sun is
    ($\odot$) is shown.
		\label{fig:XYZvtang}
	}
\end{figure*}

To measure the age of the $\delta$ Lyr cluster, we first selected a
set of candidate cluster members
(Section~\ref{sec:kinematicselection}), and then analyzed these stars
using a combination of the isochronal and gyrochronal techniques
(Section~\ref{sec:clusterage}).

\subsection{Selecting Cluster Members}
\label{sec:kinematicselection}

\citet{KounkelCovey2019} applied an unsupervised clustering algorithm
to Gaia DR2 on-sky positions, proper motions, and parallaxes for stars
within the nearest kiloparsec.  For the \cn\ (Theia 73), they reported
\noriginal\ candidate members.  We matched these stars against the
latest Gaia EDR3 observations using the \texttt{dr2\_neighbourhood}
table from the ESA archive,\footnote{See
\url{https://gea.esac.esa.int/archive/documentation/GEDR3/Gaia_archive/chap_datamodel/sec_dm_auxiliary_tables/ssec_dm_dr2_neighbourhood.html}.}
taking the stars closest in proper motion and epoch-corrected angular
distance as the presumed match \citep{gaia_collaboration_2021_edr3}.
In Figure~\ref{fig:XYZvtang}, have shown galactic positions only for
the stars with parallax signal-to-noise exceeding 20.  The reported
cluster members (gray and black points) extend over a much larger
volume than the cluster previously identified by
\citet{stephenson_possible_1959} and later corroborated by
\citet{eggen_photometric_1968}.  While the non-uniform ``clumps'' of
stars might comprise a {\it bona fide} cluster of identically-aged
stars, they could also be heavily contaminated by field stars.  We
therefore considered stars only in the immediate kinematic and spatial
vicinity of \sn\ as candidate cluster members.  We performed the
selection cuts manually, by drawing lassos with the interactive
\texttt{glue} visualization tool \citep{beaumont_2014_13866} in the
four projections shown in Figure~\ref{fig:XYZvtang}.  The overlap
between the Kepler field and the resulting candidate cluster members
is shown in Figure~\ref{fig:skychart}.  While this method will include
some field interlopers in the ``cluster star'' sample, and vice-versa,
it should suffice for our aim of verifying the existence of the
cluster in the vicinity of \sn.

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Kepler's view of the $\delta$~Lyr cluster.} Each black circle
    is a candidate cluster member selected from the catalog of
    \citet{KounkelCovey2019} based on its position and kinematics.  Of
    the \nkinematic\ candidate cluster members, 58 have at least one
    quarter of Kepler data.  TESS has also observed most of the
    cluster, for one to two lunar months to date.
		\label{fig:skychart}
	}
\end{figure}

\subsection{The Cluster's Age}
\label{sec:clusterage}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3a.pdf}
			\includegraphics[width=0.469\textwidth]{f3b.pdf}
		}

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f3c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is \clusterage\ old.}  {\it Top:} Color-absolute
    magnitude diagram of candidate \cn\ members, in addition to stars
    in IC\,2602 ($\approx38$\,Myr), the Pleiades ($\approx 115$\,Myr),
    and the Gaia EDR3 Catalog of Nearby Stars (gray background).  The
    zoomed right panel highlights the pre-main-sequence.  The \cn\ and
    IC\,2602 are approximately the same isochronal age.  {\it Bottom:}
    TESS and Kepler stellar rotation period versus dereddened Gaia
    color, with the Pleiades and Praesepe (650\,Myr) shown for
    reference \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most
    candidate \cn\ members are gyrochronally younger than the
    Pleiades; outliers are probably field interlopers.
		\label{fig:age}
	}
\end{figure*}

\subsubsection{Color-Absolute Magnitude Diagram}
\label{sec:camd}

We measured the isochrone age using an empirical approach.  The left
panel of Figure~\ref{fig:age} shows the color-absolute magnitude
diagram (CAMD) of candidate \cn\ members, IC\,2602, the Pleiades, and
the field.  The stars from the Pleiades and IC\,2602 were adopted from
\citet{CantatGaudin2018a}, and the field stars are from the Gaia EDR3
Catalog of Nearby Stars \citep{gaia_gcns_2021}.  We cleaned these
following the data filtering criteria from \citet[][Appendix
B]{GaiaCollaboration2018}, except that we weakened the parallax
precision requirement to $\varpi/\sigma_\varpi>5$.  These filters were
designed to include genuine binaries while omitting instrumental
artifacts.  We then corrected for extinction by querying the
3-dimensional maps of \citet{capitanio_threedimensional_2017} and
\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/},
2021/09/25}, and applied the extinction coefficients $k_X\equiv
A_X/A_0$ computed by \citet{GaiaCollaboration2018} assuming that $A_0
= 3.1 E(B-V)$.  For IC\,2602, the Pleiades, and the \cn, this
procedure yielded a respective mean and standard deviation for the
reddening of $E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$.
These values agree reasonably well with previously reported values
from the literature \citep[{\it
e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}.

% Gaia+2018 E(B-V):
% Pleiades 0.045, IC2602 0.031
%
% Randich+2018 E(B-V):
% Pleiades ??, IC2602 0.032 - 0.092
%
% KC2019: AV->E(B-V)
% Pleiades 0.220 AV -> 0.071
% IC2602 0.243297 -> 0.078
% Stephenson1 0.184 -> 0.059
%
% Bossini+2019 AV->E(B-V)
% PleiadesMelotte22 0.140 -> 0.045
% IC2602 0.096 -> 0.031
% Steph1: 0.157 -> 0.051

Figure~\ref{fig:age} shows that the \cn\ and IC\,2602 overlap, and
therefore are approximately the same age.  In our exploration, we also
compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
and the Upper-Centaurus-Lupus (UCL) component of the Sco OB2
association ($\approx$$16$\,Myr; \citealt{pecaut_star_2016}).  The
pre-main-sequence M dwarfs of the \cn\ were intermediate between the
latter two clusters.  To turn this heuristic interpolation into an age
measurement, we used the empirical method developed by
\citet{gagne_mutau_2020}.  In brief, we fitted the pre-main-sequence
loci of a set of reference clusters, and the locus of the target \cn\
was then modeled as a piecewise linear combination of these reference
clusters.  For our reference clusters, we adopted members of UCL,
IC\,2602, and the Pleiades from \citet{Damiani2019} and
\cite{CantatGaudin2018a} respectively.  We removed binaries by
requiring $\texttt{RUWE}<1.3$, $\texttt{radial\_velocity\_error}$
below the 80$^{\rm th}$ percentile of each cluster's distribution, and
excluded stars that were obvious photometric binaries in the CAMD.  We
then passed a moving box average and standard deviation across the
CAMD in 0.10\,mag bins, fitted a univariate spline to the binned
values, and assembled a piecewise grid of hybrid isochrones spanning
the ages between UCL to the Pleiades using Equations~6 and~7 from
\citet{gagne_mutau_2020}.

% Randich+18: 43.7 +4.3 - 3.9 LDB
% Dobbie+10: 46 +6 -5 LDB

The ages returned by this procedure depend on the ages assumed for
each reference cluster.  We adopted a 115\,Myr age for the Pleiades
\citep{dahm_2015}, and a 16\,Myr age for UCL \citep{pecaut_star_2016}.
The age of IC\,2602 however is the most important ingredient, since it
receives the most weight in the interpolation.  Plausible ages for
IC\,2602 span 30\,Myr to 46\,Myr, with older ages being preferred by
the lithium-depletion-boundary (LDB) measurements
\citep{dobbie_ic_2010,randich_gaiaeso_2018} and younger ages by the
main-sequence turn-off
\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.  If
we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would be
$31^{+5}_{-4}$\,Myr old.  For the converse extreme of 46\,Myr, the
\cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
38\,Myr age for IC\,2602, which yields an age for the \cn\ of
\clusterage.\footnote{Our exploration of the PARSEC and MIST isochrone
models over a grid of ages, metallicities, and reddenings, yielded the
best agreement for this $\approx 38\,$Myr age as well, given ${\rm
[Fe/H] = +0.1}$ and $A_V=0.2$
\citep{bressan_parsec_2012,choi_mesa_2016};  this preferred CAMD
reddening is higher than the \citet{lallement_threedimensional_2018}
value by a factor of two.  } Follow-up studies of the LDB or
main-sequence turn-off in the \cn\ could help determine a more precise
and accurate age for the cluster, and are left for future work.

\subsubsection{Stellar Rotation Periods}

Of the \noriginal\ candidate \cn\ members reported by
\citet{KounkelCovey2019}, \nwithtess\ stars were amenable to rotation
period measurements ($G<17$ and $\bpmrpo>0.5$) using the TESS full
frame image data.\footnote{Kepler rotation periods were derived by
\citet{mcquillan_rotation_2014} for 28 stars; for these cases, we
adopted the Kepler rotation period.  } We extracted light curves from
the TESS images using the nearest pixel to each star, and regressed
them against systematics with the causal pixel model implemented in
the \texttt{unpopular} package \citep{hattorio_2021_cpm}.  We then
measured candidate rotation periods using a Lomb-Scargle periodogram
\citep{lomb_1976,scargle_studies_1982,astropy_2018}.  To enable cuts
on crowding, we queried the Gaia source catalog for stars within a
$21\farcs0$ radius of the target star (a radius of 1 TESS pixel).
Within this radius, we recorded the number of stars with greater
brightness than the target star, and with brightness within 1.25 and
2.5 TESS magnitudes of the target star.

We then cleaned the candidate TESS rotation period measurements
through a combination of automated and manual steps.  As a matter of
scope, we restricted our attention to the \nkinwithtess\ stars
discussed in Section~\ref{sec:kinematicselection} in the spatial and
kinematic proximity of Kepler\,1627.  To avoid blending, we also only
focused on stars for which no companions were known with a brightness
exceeding one-tenth of the target star in a $21\farcs0$ radius.  There
were \nkinwithtessandcrowding\ stars that met these requirements, and
that had TESS data available.  For plotting purposes we then imposed a
selection based on the strength of the signal itself: we required the
Lomb Scargle power to exceed 0.2, and the period to be below 15\,days.

The lower panel of Figure~\ref{fig:age} shows the resulting
\nkindefaultcleaning\ stars.  The majority of these stars fall below
the ``slow sequence'' of the Pleiades, consistent with a gyrochronal
age for the \cn\ below 100\,Myr.  In fact, the rotation-color
distributions of other 30\,Myr to 50\,Myr clusters ({\it e.g.},
IC\,2602 and IC\,2391) are indistinguishable
\citep{douglas_stephanie_t_2021_5131306}.  Approximately 10 of the
\cn\ stars appear as outliers above the ``slow sequence''.  Assuming
that they are all false positives ({\it i.e.}, field interlopers), our
rotation period detection fraction would be
$\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.   The other
stars are likely to be field contaminants.  A final possible
confounding factor -- binarity -- is known to affect the ``fast
sequence'' of stars beneath the slow sequence
\citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.  We do
not expect it to change the central conclusion regarding the cluster's
age.

\section{The Stars}
\label{sec:stars}

\subsection{Kepler\,1627A}
\subsubsection{Age}
% rotation period from plot_rotation_period_windowslider.py
Based on the spatial and kinematic association of \sn\ with the \cn,
and the assumption that the planet formed shortly after the star, it
seems likely that \sn\ is the same age as the cluster. There are two
consistency checks on whether this is true: rotation and lithium.
Based on the Kepler light curve, the rotation period is
$2.642\pm0.042$\,days, where the quoted uncertainty is based on the
scatter in rotation periods measured from each individual Kepler
quarter.  This is consistent with comparable cluster members
(Figure~\ref{fig:age}).

To infer the amount of Li\,\textsc{I} from the 6708\,\AA\ doublet
\citep[{\it e.g.},][]{soderblom_ages_2014}, we acquired an iodine-free
spectrum from Keck/HIRES on the night of 2021 March 26 using the
standard setup and reduction techniques of the California Planet
Survey \citep{howard_cps_2010}.  Following the equivalent width
measurement procedure described by \citet{bouma_2021_ngc2516}, we find
${\rm EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   This value does not
correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Nonetheless,
given the stellar effective temperature (Table~\ref{tab:starparams}),
this measurement is in agreement with expectations for a
$\approx40$\,Myr star ({\it e.g.}, as measured in IC~2602 by
\citealt{randich_gaiaeso_2018}).  It is also larger than any lithium
equivalent widths measured by \citet{berger_identifying_2018} in their
analysis of 1{,}301 Kepler-star spectra.

\subsubsection{Stellar Properties}

The adopted stellar parameters are listed in
Table~\ref{tab:starparams}.  The stellar mass, radius, and effective
temperature are found by interpolating against a 38$\,$Myr MIST
isochrone \citep{choi_mesa_2016}.  The statistical uncertainties are
propagated from the absolute magnitude (mostly originating from the
parallax uncertainty) and the color; the systematic uncertainties are
taken to be the difference between the PARSEC
\citep{bressan_parsec_2012} and MIST isochrones.  Reported
uncertainties are a quadrature sum of the statistical and systematic
components.  As a consistency check, we analyzed the aforementioned
Keck/HIRES spectrum from the night of 2021 March 26 using a
combination of \texttt{SpecMatch-Emp} for stellar properties, and
\texttt{SpecMatch-Synth} for $v\sin i$ \citep{yee_SM_2017}.  This
procedure yielded $T_{\rm eff}=5498\pm100\,{\rm K}$, $\log
g=4.6\pm0.1$, $[{\rm Fe/H}]=0.15\pm0.10$ from \texttt{SpecMatch-Emp},
and $v\sin i = 18.9\pm1.0$ from \texttt{SpecMatch-Synth}.  These
values are within the $1$-$\sigma$ uncertainties of our adopted values
from the isochrone interpolation.

\subsection{Kepler\,1627B}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f4a.pdf}
			\includegraphics[width=0.48\textwidth]{f4b.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2.5V
    companion at $\rho \approx 0\farcs16$, which corresponds to a
    projected separation of $53\pm4$\,AU.  The inset shows a cutout of
    the stacked NIRC2 image (North is up, East is left, scale is set
    by the separation of the binary).  The lines show 5-$\sigma$
    contrast limits for the `Alopeke filters, and 6-$\sigma$ contrast
    limits for NIRC2 outside of $0\farcs15$. {\it Right:} Gaia EDR3
    renormalized unit weight error (RUWE) point estimates for
    candidate $\delta$\,Lyr cluster members.  Since other members of
    the cluster with similar brightnesses have comparable degrees of
    photometric variability, the high RUWE independently suggests that
    Kepler\,1627 is a binary. 
    \label{fig:kep1627binary}
	}
\end{figure*}

We first noted the presence of a close neighbor in the \sn\ system on
2015 July 22 when we acquired adaptive optics imaging using the NIRC2
imager on Keck-II.  We used the narrow camera (FOV = 10.2\arcsec) to
obtain 8 images in the $K'$ filter ($\lambda = 2.12\,\mu$m) with a
total exposure time of 160\,s. We analyzed these data following
\citet{kraus_impact_2016}, which entailed using PSF-fitting to measure
the separation, position angle, and contrast of the candidate
companion.  The best-fitting empirical PSF template was identified
from among the near-contemporaneous observations of single stars in
the same filter.  The mean values inferred from the 8 images are
reported in Table~\ref{tab:starparams}.  To estimate the detection
limits, we analyzed the residuals after subtracting the empirical PSF
template. Within each residual image, the flux was measured through
40\,mas apertures centered on every pixel, and then the noise as a
function of radius was estimated from the RMS within concentric rings.
Finally, the detection limits were estimated from the strehl-weighted
sum of the detection significances in the image stack, and we adopted
the $6$-$\sigma$ threshold as the detection limit for ruling out
additional companions.

We also observed \sn\ on Gemini-North using the `Alopeke speckle
imager on 2021 June 24.  `Alopeke is a dual-channel speckle
interferometer that uses narrow-band filters centered at 0.83\,$\mu$m
and $0.56\,\mu$m.  We acquired three sets of $1000\times 60$$\,$msec
exposures during good seeing (0.45$''$), and used the autocorrelation
function of these images to reconstruct a single image and 5-$\sigma$
detection limits (see \citealt{howell_speckle_2011}).  This procedure
yielded a detection of the companion in the 0.83\,$\mu$m notch filter,
but not the $0.56\,\mu$m filter.  The measured projected separation
and magnitude difference are given in Table~\ref{tab:starparams}.

% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
Figure~\ref{fig:kep1627binary} summarizes the results of the
high-resolution imaging.  The Gaia EDR3 parallax for the primary
implies a projected separation of $53 \pm 4$\,AU, assuming the
companion is bound.  Although the companion is unresolved in the Gaia
source catalog (there are no comoving, codistant candidate companions
brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its
existence was also suggested by the primary star's large renormalized
unit weight error (RUWE), relative to other members of the \cn.  Based
on the apparent separation, the binary orbital period is of order
hundreds of years.  The large RUWE is therefore more likely to be
caused by a PSF-mismatch skewing the Gaia centroiding during
successive scans, rather than true astrometric motion.  Regardless,
given the low geometric probability that a companion imaged at $\rho
\approx 0\farcs16$ is a chance line-of-sight companion, we proceed
under the assumption that the companion is bound, and that
Kepler\,1627 is a binary.  Given the distance and age, the models of
\citet{baraffe_new_2015} imply a companion mass of $M_{\rm B} \approx
0.33 M_{\odot}$ and companion temperature of $T_{\rm eff,B} \approx
3450$ K.  The corresponding spectral type is roughly M2.5V
\citep{pecaut_mamajek_2013}.  These models combined with the NIRC2
contrast limits imply physical limits on tertiary companions of
$M_{\rm ter} < 50 M_{\rm Jup}$ at $\rho = 50$\,AU, $M_{\rm ter} < 20
M_{\rm Jup}$ at $\rho = 100$\,AU, and $M_{\rm ter} < 10 M_{\rm Jup}$
at $\rho = 330$\,AU.

\section{The Planet}
\label{sec:planet}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=\textwidth]{f5a.pdf}
		}

		\vspace{-0.2cm}	
		\subfloat{
			\includegraphics[width=0.5\textwidth]{f5b.pdf}
			\includegraphics[width=0.5\textwidth]{f5c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
    The Kepler data span 1{,}437 days (3.9 years), sampled at 30
    minute cadence;  a 100 day segment is shown.  The top panel shows
    the \texttt{PDCSAP} median-subtracted flux in units of
    parts-per-thousand ($\times 10^{-3}$).  The dominant signal is
    induced by starspots.  The stellar variability model (orange line)
    is subtracted below, revealing the transits of \pn.  The online
    Figure Set spans the entire 3.9 years of observations.
    {\it Bottom}:
    Phase-folded transit of Kepler 1627Ab with stellar variability
    removed.  Windows over 20 hours ({\it left}) and the entire orbit
    ({\it right}) are shown, and the residual after subtracting the
    transit is in the bottom-most row.  The 2-$\sigma$ model uncertainties
    and the best-fit model are the light purple band and the dark
    purple line.  Gray points are individual flux measurements; black
    points bin these to 20 minute intervals, and have a representative
    1-$\sigma$ error bar in the lower right of each panel.  The asymmetric
    residual during transit is larger than the out-of-transit scatter.
    \label{fig:lc}
  }
\end{figure*}

\subsection{Kepler Light Curve}

The Kepler space telescope observed \sn\ at a 30-minute cadence from
2009 May 2 until 2013 April 8.  Data gaps during quarters 4, 9, and 13
led to an average duty cycle over the 3.9~year interval of 67\%.  \sn\
was also observed at 1-minute cadence from 2012 Oct 5 until 2013 Jan
11.  The top panel of Figure~\ref{fig:lc} shows a portion of the
30-minute cadence \texttt{PDCSAP} light curve.  Nonastrophysical
variability has been removed using the methods discussed by
\citet{smith_kepler_PDC_2017}; the default optimal aperture was
assumed \citep{smith_finding_2016}.  Cadences with non-zero quality
flags (9\% of the data) have been omitted.  The resulting photometry
is dominated by a quasi-periodic starspot signal with a peak-to-peak
amplitude that varies between 2\% and 8\%.  Previous analyses have
identified and characterized the smaller transit signal
\citep{2012ApJS..199...24T,thompson_planetary_2018}, validated its
planetary nature \citep{morton_false_2016}, and even searched the
system for transit timing variations \citep{holczer_transit_2016}.
Nonetheless, since the cluster membership provides us with more
precise stellar parameters than those previously available, we opted
to reanalyze the light curve.

We fitted the Kepler long cadence time series with a model that
simultaneously included the planetary transit and the stellar
variability.  The stellar variability was modeled with the
\texttt{RotationTerm} Gaussian Process kernel in \texttt{exoplanet}
\citep{exoplanet:exoplanet}.  This kernel assumes that the variability
is generated by a mixture of two damped simple harmonic oscillators
with characteristic frequencies set by 1/$P_{\rm rot}$ and its first
harmonic.  We additionally included a jitter term to inflate the flux
uncertainties in a manner that accounted for otherwise unmodeled
excess white noise, and let the eccentricity float.  For the
limb-darkening, we assumed a quadratic law, and sampled using the
uninformative prior suggested by \citet{exoplanet:kipping13}.

Our model therefore included 10 free parameters for the transit ($\{P,
t_0, \delta, b, u_1 ,u_2 ,R_\star, \log g, e, \omega \}$), 2 free
parameters for the light curve normalization and a white noise jitter
($\{\langle f \rangle, \sigma_f \}$), and 5 hyperparameters for the GP
($\{\sigma_{\mathrm{rot}}, P_{\mathrm{rot}}, Q_0, \mathrm{d}Q, f \}$).
We also considered including an additive \texttt{SHOTerm} kernel to
account for stochastic noise, but found that this did not affect the
results, and so opted for the simpler GP kernel.  We fitted the models
using \texttt{PyMC3} \citep{salvatier_2016_PyMC3,exoplanet:theano},
and accounted for the finite integration time of each exposure in the
numerical integration when evaluating the model light curve
\citep[see][]{kipping_binning_2010}.  We assumed a Gaussian
likelihood, and after initializing each model with the parameters of
the maximum {\it a posteriori} model, we sampled using
\texttt{PyMC3}'s gradient-based No-U-Turn Sampler
\citep{hoffman_no-u-turn_2014} in the bases indicated in
Table~\ref{tab:posterior}.  We used $\hat{R}$ as our convergence
diagnostic \citep{gelman_inference_1992}.

Figure~\ref{fig:lc} shows the resulting best-fit model in orange (top)
and purple (bottom).  The model parameters and their uncertainties,
given in Table~\ref{tab:posterior}, are broadly consistent with a
mini-Neptune sized planet ($3.82\pm0.16\,R_\oplus$) on a close-in
circular\footnote{ Our transit fitting yields $e<0.48$ at 2-$\sigma$;
the constraints on the eccentricity are not particularly strong.}
orbit around a G8V host star ($0.88 \pm 0.02 R_\odot$).  This best-fit
planet size is consistent with those previously reported by
\citet{morton_false_2016} and \citet{berger_identifying_2018}, and
corrects for the small amount of flux dilution from Kepler\,1627B.

The transit fit however is not perfect: the lower panels of
Figure~\ref{fig:lc} show an asymmetric residual in the data relative
to the model: the measured flux is high during the first half of
transit, and low in the second half.  The semi-amplitude of this
deviation is $\approx50\,{\rm ppm}$, which represents a $\approx 3\%$
distortion of the transit depth ($1759\pm62\,{\rm ppm}$).  Note that
while this asymmetry is within the 2-$\sigma$ model uncertainties, the
model has a jitter term that grows to account for otherwise unmodeled
white noise in the flux.  The significance of the asymmetry is
therefore best assessed in comparison against the intrinsic
out-of-transit scatter in the data, not the model uncertainties.

To determine whether the asymmetry is a systematic caused by our
stellar variability model, we explored an alternative approach in
which we isolated each transit window and locally fitted out
polynomial trends, and then binned all the observed transits; the
asymmetry was still present at a comparable amplitude.
Appendix~\ref{app:asymmetry} includes a deeper analysis, and finds
that the asymmetry also seems to be robust to different methods of
data binning in time and by local light curve slope.  Possible
astrophysical explanations are discussed in Section~\ref{sec:conc}.

\subsection{Planet Confirmation}

\begin{figure}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.49\textwidth]{f6.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Evidence for a prograde orbit of Kepler 1627 Ab.} The time of
    each Kepler transit was measured, along with the local slope of
    the light curve.  The two quantities appear anti-correlated, which
    is most easily explained by starspot crossings during the first
    (second) half of transit inducing a positive (negative) TTV,
    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
    units along the abscissa can be understood by considering that the
    stellar flux changes by $\sim$60\,ppt per half rotation period
    ($\sim$1.3\,days).
		\label{fig:corr}
	}
\end{figure}

If the \pn\ transit signal is created by a genuine planet,  then to
our knowledge it would be the youngest planet yet found by the main
Kepler mission.\footnote{The re-purposed K2 mission however has found
two younger systems containing five planets: K2-33b ($9\pm1\,{\rm
Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}   Could the transit
be produced by anything other than a planet orbiting this near-solar
analog?  \citet{morton_false_2016} validated the planet based on the
transit shape, arguing that the most probable false positive scenario
was that of a background eclipsing binary, which had a model-dependent
probability of $\approx$$10^{-5}$.  However, this calculation was
performed without knowledge of the low-mass stellar companion ($M_{\rm
B}\approx 0.33\,M_\odot$).  Validated planets have also previously
been refuted \citep[{\it e.g.},][]{shporer_three_2017}.  We therefore
reassessed false positive scenarios in some detail.

% (0.7-2.0\% V-band to G-band dmag difference from EEM table)

As an initial plausibility check, Kepler\,1627B contributes 1\% to 2\%
of the total flux observed in the Kepler aperture.  For the sake of
argument, assume the former value.  The observed transit has a depth
of $\approx$0.18\%.  A 18\% deep eclipse of Kepler\,1627B would
therefore be needed to produce a signal with the appropriate depth.
The shape of the transit signal however requires the impact parameter
to be below 0.77 (2-$\sigma$); the tertiary transiting the secondary
would therefore need to be non-grazing with $R_3/R_2 \approx 0.4$.
This yields a contradiction:  this scenario would require an ingress
and egress phase that each span $\approx$40\% of the transit duration
($\approx 68\,{\rm minutes}$).  The actual measured ingress and egress
duration is $\approx 17\,{\rm minutes}$, a factor of four times too
short.  The combination of Kepler 1627B's brightness, the transit
depth, and the ingress duration therefore disfavor the scenario that
Kepler 1627B might host the transit signal.

% Baraffe+15: M_B 0.33 Mdot
% Rstar = 0.455 --> 0.465 Rsun

There are two other new lines of evidence that confirm the planetary
interpretation.  First, the stellar density implied by the transit duration and orbital period
is inconsistent with an eclipsing body around the M-dwarf companion:
we find $\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
theoretically expected density for Kepler\,1627B is $\approx 4.6\,{\rm
g\,cm}^{-3}$ (Table~\ref{tab:posterior}; \citealt{choi_mesa_2016}).
The transit duration is therefore too long to be explained by a star
eclipsing the M dwarf secondary at $10$-$\sigma$.  While the planet
might hypothetically still orbit a hidden close and bright companion,
this possibility is ruled out by {\it i)} a lack of secondary lines in
the HIRES spectra and {\it ii)} a lack of secondary rotation signals
in the Kepler photometry.

%NOTE to the arxiv readers:, the $\approx$100 days of short-cadence
%Kepler observations also mitigate the cadence-induced smearing, and
%enable a more precise measurement of the transit impact parameter.

The second line of confirming evidence comes by analyzing the
individual Kepler transit times. We isolated each of the 144 observed
transits to within $\pm4.5$\,hr of each transit, and fitted each
window with both {\it i)} a local second-order polynomial and transit,
and {\it ii)} a local linear trend.  We let the mid-time of each
transit float, and then calculated the residual between the measured
mid-time and that of a periodic orbit.  This residual, the transit
timing variation (TTV), is plotted in Figure~\ref{fig:corr} against
the local linear slope.  A significant correlation of $-0.0564 \pm
0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
Objects of Interest have shown this correlation
\citep{holczer_time_2015}, which is most readily interpreted as a TTV
induced by unresolved starspot crossings \citep{mazeh_time_2015}.
This is only possible if the planet transits the primary star, which
excludes a background eclipsing binary scenario, and therefore
confirms that \pn\ is a planet.  It also suggests that the planet's
orbit is prograde.  The latter point assumes that the dominant
photometric variability is induced by dark spots, and not bright
faculae.  Given the observed transition of Sun-like stellar
variability from spot to faculae-dominated regimes between young and
old ages, we expect this assumption to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.49\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Radial velocities acquired using Keck/HIRES during the
    transit of 2021 Aug 7, observed simultaneously with MuSCAT3.}
    Shaded bands show 2-$\sigma$ model uncertainties.  The photometric
    transit depths are consistent across the {\it griz} bandpasses.
    The best-fitting model in the top panel includes the RM effect and
    a quadratic trend in time to model the spot-induced
    $\approx$\,$250$\ms\ trend (see Appendix~\ref{app:rm}).  The
    radial velocity jitter over the 15~minute exposure timescale
    ($\sigma_{\rm RV}\approx30$\ms) prevented us from detecting the RM
    effect.  The RV variations are strongly correlated with varying
    emission in the Ca H and K lines (middle panel).
		\label{fig:ground}
	}
\end{figure}

A third supporting line of evidence for the planetary interpretation
exists, though we consider it less definitive than the stellar density
and TTV-local slope correlation.  We observed a transit of \pn\ on the
night of 2021 Aug 7 simultaneously with Keck/HIRES and MuSCAT3.  We
scheduled the observations using the ephemeris of
\citet{holczer_transit_2016}.  Although we did not detect the
Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
show the transit (Figure~\ref{fig:ground}).  Fitting the MuSCAT3
photometry with a model that lets the transit depths vary across each
bandpass, we find {\it griz} depths consistent with the Kepler depth
at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  Conditioned on the
Kepler ephemeris and transit depth, the MuSCAT3 observations may also
show a transit duration $17.3\pm4.0\ {\rm minutes}$ shorter than the
Kepler transits; further photometric follow-up would help confirm
whether the transit duration is indeed changing.  For our RM analysis,
the details are discussed in Appendix~\ref{app:rm}.  While the
velocities are marginally more consistent with a prograde or polar
orbit than a retrograde orbit, the spot-corrected exposure-to-exposure
scatter ($\sigma_{\rm RV}\approx 30$\ms) is comparable to the expected
RM anomaly assuming an aligned orbit ($\Delta v_{\rm RM}\approx
20\,{\rm m\,s}^{-1}$).  We are therefore not in a position to claim a
spectroscopic detection of the RM effect, nor to quantify the stellar
obliquity.

\section{Discussion \& Conclusions}
\label{sec:conc}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.74\textwidth]{f8.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
    3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
    its uncertainty.  \pn\ is shown with a star.  The large sizes of
    the youngest transiting planets could be explained by their
    primordial atmospheres not yet having evaporated; direct
    measurements of the atmospheric outflows or planetary masses would
    help to confirm this expectation.  Selection effects may also be
    important.  Parameters are from the NASA Exoplanet Archive (2021
    Sept 15).
		\label{fig:rp_period_age}
	}
\end{figure*}

Kepler\,1627Ab provides a new extremum in the ages of the Kepler
planets, and opens multiple avenues for further study.  Observations
of spectroscopic transits at greater precision should yield a
measurement of the stellar obliquity, which would confirm or refute
the prograde orbital geometry suggested by the TTV-local slope
correlation.  Separately, transit spectroscopy aimed at detecting
atmospheric outflows could yield insight into the evolutionary state
of the atmosphere \citep[{\it
e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
Observations aimed at quantifying the amount of high-energy
irradiation incident on the planet would complement these efforts, by
helping to clarify the expected outflow rate \citep[{\it
e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
quantity to measure would be the planet's mass.  The most feasible
approach would likely be long-term monitoring of the transit times and
durations, given the technical difficulty of performing a multi-color
radial velocity campaign on a $V=13$ star.  Measured at
sufficient precision though, the mass, combined with the known age and
size, would yield constraints on both the planet's composition and its
initial entropy \citep{owen_constraining_2020}.

More immediately, the Kepler data may yet contain additional
information.  For instance, one possible explanation for the transit
asymmetry shown in Figure~\ref{fig:lc} is that of a dusty asymmetric
outflow.  A second possibility is that the planetary orbit is slightly
misaligned from the stellar spin axis, and tends to transit starspot
groups at favored stellar latitudes.  Other possibilities such as
gravity darkening or TTVs causing the asymmetry are disfavored (see
Appendix~\ref{app:asymmetry}).  Dusty outflows are theoretically
expected for young mini-Neptunes, and the amplitude of the observed
asymmetry is also consistent with predictions \citep{wang_dai_2019}.
Beyond the asymmetric transits, Appendix~\ref{app:flare} highlights an
additional abnormality in the Kepler data, in the arrival time
distribution of stellar flares.  We encourage its exploration by
investigators more versed in the topic than ourselves.

In the context of the transiting planet population, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precise ages include K2-33
\citep{Mann_K2_33b_2016,David_et_al_2017}, DS~Tuc
\citep{benatti_possibly_2019,newton_tess_2019} , HIP~67522
\citep{rizzuto_tess_2020}, TOI~837 \citep{bouma_cluster_2020}, the
two-planet AU~Mic \citep{plavchan_planet_2020,martioli_aumicbc_2021}
and the four-planet V1298~Tau \citep{david_four_2019}.  \pn\ is one of
the smaller planets in this sample ($3.82\pm0.16\,R_\oplus$), which
could be linked to the selection effects imposed by spot-induced
photometric variability at very young ages \citep[{\it
e.g.},][]{zhou_2021_tois}.  However, smaller planets could have been
detected: the Kepler pipeline's median completeness extended to
$1.6\,R_\oplus$ at 10~day orbital periods, and $3.3\,R_\oplus$ at
100~days \citep{2021ascl.soft07027B}.  The large size of \pn\ relative
to most Kepler mini-Neptunes might therefore support a picture
in which the typical $5\,M_\oplus$ to 10$\,M_\oplus$ mini-Neptune
\citep{wu_mass_2019} loses a significant fraction of its primordial
atmosphere over its first gigayear
\citep{Owen_Wu_2013,ginzburg_corepowered_2018}.

% The prograde orbit, if confirmed, would be consistent with a
% quiescent disk-driven migration history, and would differ from
% misalignments that have been observed for some older hot
% Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
%Spectroscopic follow-up observations of the transit remain necessary
%however to verify whether the TTV-slope correlation is yielding the
%correct orbital geometry.  The alternative scenario of in-situ
%formation is disfavored on theoretical grounds: the required disk
%mass would likely drive migration, and so lacks internal consistency
%\citep{inamdar_formation_2015,ogihara_reassessment_2015}.

Ultimately, the main advance of this work is a precise measurement of
the age of \pn.   This measurement was enabled by identifying the
connection of the star to the \cn\ using Gaia kinematics, and then by
using the Gaia color-absolute magnitude diagram and TESS stellar
rotation periods to verify the cluster's existence.
Table~\ref{tab:v06} enables similar cross-matches for both known and
forthcoming exoplanet systems \citep[{\it
e.g.},][]{guerrero_tess_2021}. Confirming these candidate associations
using independent age indicators is essential because their false
positive rates are not known.  A related path is to identify new
kinematic associations around known exoplanet host stars using
positions and tangential velocities from Gaia, and to verify
these associations with stellar rotation periods and spectroscopy
\citep[{\it e.g.},][]{tofflemire_tess_2021}.  Each approach seems
likely to expand the census of planets with precisely measured ages
over the coming years, which will help in deciphering the early stages
of exoplanet evolution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
\acknowledgements
\raggedbottom

The authors are grateful to J{.}~Winn, J{.}~Spake, A{.}~Howard, and
T{.}~David for illuminating discussions and suggestions, and to
R{.}~Kerr for providing us with the \citet{Kerr2021} membership list
prior to its publication.  The authors are also grateful to
K{.}~Collins for helping resolve the scheduling conflict that would
have otherwise prevented the MuSCAT3 observations.
%
L.G.B{.} acknowledges support from a Charlotte Elizabeth Procter
Fellowship from Princeton University, as well as from the TESS GI
Program, programs G011103 and G022117, through NASA grants
80NSSC19K0386 and 80NSSC19K1728.
%
%
Keck/NIRC2 imaging was acquired by program 2015A/N301N2L
(PI: A.~Kraus). % and 2019A/N069 (PI: E.~Petigura).
%
In addition, this paper is based in part on observations made with the
MuSCAT3 instrument, developed by the Astrobiology Center and under
financial support by JSPS KAKENHI (JP18H05439) and JST PRESTO
(JPMJPR1775), at Faulkes Telescope North on Maui, HI, operated by the
Las Cumbres Observatory.
%
This work is partly supported by JSPS KAKENHI Grant Numbers 22000005,
JP15H02063, JP17H04574, JP18H05439, JP18H05442, JST PRESTO Grant
Number JPMJPR1775, the Astrobiology Center of National Institutes of
Natural Sciences (NINS) (Grant Number AB031010).
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%
%
%This study was based in part on observations at Cerro Tololo
%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
%Association of Universities for Research in Astronomy (AURA) under a
%cooperative agreement with the National Science Foundation.
%
%
Finally, this research has made use of the Keck Observatory Archive (KOA),
which is operated by the W. M. Keck Observatory and the NASA Exoplanet
Science Institute (NExScI), under contract with the National
Aeronautics and Space Administration.  We also thank the Keck
Observatory staff for their support of HIRES and remote observing.  We
recognize the importance that the summit of Maunakea has within the
indigenous Hawaiian community, and are deeply grateful to have the
opportunity to conduct observations from this mountain.
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

% %
% This research made use of the NASA Exoplanet Archive, which is
% operated by the California Institute of Technology, under contract
% with the National Aeronautics and Space Administration under the
% Exoplanet Exploration Program.
% %

% Resources supporting this work were provided by the NASA High-End
% Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
% Division at Ames Research Center for the production of the SPOC data
% products.
%

% A.J.\ and R.B.\ acknowledge support from project IC120009 ``Millennium
% Institute of Astrophysics (MAS)'' of the Millenium Science Initiative,
% Chilean Ministry of Economy. A.J.\ acknowledges additional support
% from FONDECYT project 1171208.  J.I.V\ acknowledges support from
% CONICYT-PFCHA/Doctorado Nacional-21191829.  R.B.\ acknowledges support
% from FONDECYT Post-doctoral Fellowship Project 3180246.
% %
% C.T.\ and C.B\ acknowledge support from Australian Research Council
% grants LE150100087, LE160100014, LE180100165, DP170103491 and
% DP190103688.
% %
% C.Z.\ is supported by a Dunlap Fellowship at the Dunlap Institute for
% Astronomy \& Astrophysics, funded through an endowment established by
% the Dunlap family and the University of Toronto.
% %
% D.D.\ acknowledges support through the TESS Guest Investigator Program
% Grant 80NSSC19K1727.
%
%
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

\software{
  %\texttt{arviz} \citep{arviz_2019},
  \texttt{altaipony} \citep{ilin_flares_2021},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  %\texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
   	exoplanet:theano},
	%\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  %\texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  %\texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  %\texttt{numpy} \citep{walt_numpy_2011}, 
  %\texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  \texttt{TESS-point}  \citep{burke_2020},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
\

\facilities{
 	{\it Astrometry}:
 	Gaia \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2021_edr3}.
 	{\it Imaging}:
    Second Generation Digitized Sky Survey. %,
    %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
 	Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	Gemini:North~(`Alopeke; \citealt{scott_nessi_2018,scott_twin_2021}.
 	{\it Spectroscopy}:
	%CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
  %PFS ({\bf CITE}),
  %  MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
	%AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
 	Keck:I~(HIRES; \citealt{vogt_hires_1994}).
 	%VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
%	  ASTEP:0.40$\,$m (ASTEP400),
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
%	  El Sauce:0.356$\,$m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
	  Kepler \citep{borucki_kepler_2010},
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
    MuSCAT3 \citep{Narita_2020},
% 	NTT,
% 	SOAR (SOI),
 	  TESS \citep{ricker_transiting_2015}.
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
}

\input{starparams.tex}
\input{model_posterior_table.tex}
\input{t1_appendix.tex}
\input{t2_appendix.tex}

\clearpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\appendix
\section{Young, Age-Dated, and Age-Dateable Star Compilation}
\label{app:targetlist}

The \texttt{v0.6} CDIPS target catalog (Table~\ref{tab:v06}) includes
stars that are young, age-dated, and age-dateable.  By
``age-dateable'', we mean that the stellar age should be measurable at
greater precision than that of a typical FGK field star, through
either isochronal, gyrochronal, or spectroscopic techniques.  As in
\citet{bouma_cdipsI_2019}, we collected stars that met these criteria
from across the literature.  Table~\ref{tab:metadata} gives a list of
the studies included, and brief summary statistics.  The age
measurement methodologies adopted by each study differ: in many,
spatial and kinematic clustering has been performed on the Gaia data,
and ensemble isochrone fitting of the resulting clusters has been
performed (typically focusing on the turn-off).  In other studies
however, the claim of youth is based on the location of a single star
in the color-absolute magnitude diagram, or on spectroscopic
information.

One major change in Table~\ref{tab:v06} relative to the earlier
iteration from \citet{bouma_cdipsI_2019} is that the extent of
Gaia-based analyses has now matured to the point that we can neglect
pre-Gaia cluster memberships, except for a few cases with
spectroscopically confirmed samples of age-dated stars.  The
membership lists for instance of \citet{Kharchenko_et_al_2013} and
\citet{dias_proper_2014} (MWSC and DAML) are no longer required.  This
is helpful for various post-processing projects,  since the field star
contamination rates were typically much higher in these catalogs than
in the newer Gaia-based catalogs.

The most crucial parameters of a given star for our purposes are the
Gaia DR2 $\texttt{source\_id}$, the cluster or group name
($\texttt{cluster}$), and the $\texttt{age}$.  Given the hierarchical
nature of many stellar associations, we do not attempt to resolve the
cluster names to a single unique string.  The Orion complex for
instance can be divided into almost one hundred kinematic subgroups
\citep{kounkel_apogee2_2018}.  Similar complexity applies to the
problem of determining homogeneous ages, which we do not attempt to
resolve.  Instead, we simply merged the cluster names and ages
reported by various authors into a comma-separated string.

This means that the \texttt{age} column can be null, for cases in
which the original authors did not report an age, or for which a
reference literature age was not readily available.  Nonetheless,
since we do prefer stars with known ages, we made a few additional
efforts to populate this column.  When available, the age provenance
is from the original analysis of the cluster.  In a few cases however
we adopted other ages when string-based cross-matching on the cluster
name was straightforward.  In particular, we used the ages determined
by \citet{CantatGaudin2020b} to assign ages to the clusters from
\citet{GaiaCollaboration2018}, \citet{CantatGaudin2018a},
\citet{CastroGinard2020}, and \citet{CantatGaudin2020a}.

The catalogs we included for which ages were not immediately available
were those of \citet{CottenSong2016}, \citet{Oh2017},
\citet{Zari2018}, \citet{Gagne2018b}, \citet{Gagne2018a},
\citet{Gagne2018c}, and \citet{Ujjwal2020}.  While in principle the
moving group members discussed by
\citet{Gagne2018b,Gagne2018a,Gagne2018c} and \citet{Ujjwal2020} have
easily associated ages, our SIMBAD cross-match did not retain the
moving group identifiers given by those studies, which should
therefore be recovered using tools such as BANYAN
$\Sigma$.\footnote{\url{http://www.exoplanetes.umontreal.ca/banyan/banyansigma.php}}.
We also included the SIMBAD object identifiers $\texttt{TT*}$,
$\texttt{Y*O}, $\texttt{Y*?}, $\texttt{TT?}$, and $\texttt{pMS*}$.
Finally, we  included every star in the NASA Exoplanet Archive
planetary system ($\texttt{ps}$) table that had a Gaia identifier
available \citep{NASAExoArchive_ps_20210506}.  If the age had finite
uncertainties, we also included it, since stellar ages determined
through the combination of isochrone-fitting and transit-derived
stellar densities typically have higher precision than from isochrones
alone.

For any of the catalogs for which Gaia DR2 identifiers were not
available, we either followed the spatial (plus proper-motion)
cross-matching procedures described in \citet{bouma_cdipsI_2019}, or
else we pulled the Gaia DR2 source identifiers associated with the
catalog from SIMBAD.  We consequently opted to drop the
$\texttt{ext\_catalog\_name}$ and $\texttt{dist}$ columns maintained
in \citet{bouma_cdipsI_2019}, as these were only populated for a small
number of stars.  The technical manipulations for the merging,
cleaning, and joining were performed using $\texttt{pandas}$
\citep{mckinney-proc-scipy-2010}.  The eventual cross-match (using the
Gaia DR2 $\texttt{source\_id}$) against the Gaia DR2 archive was
performed asychronously on the Gaia archive
website.\footnote{\url{https://gea.esac.esa.int/archive/}}

% \section{Binarity and Stellar Rotation for $\delta$\,Lyr Cluster Members}
% \label{app:rotationbinarity}
% 
% \begin{figure}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.7\textwidth]{f7.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
%   \caption{ {\bf Binarity indicators for rotators in \cn.} 
%     Data are as in the lower panel of Figure~\ref{fig:age}.  Sources
%     with $\mathrm{RUWE}>1.2$ appear in red; these include astrometric binaries as well
%     as sources poorly fit by a single-star PSF.  Photometric binaries
%     ($>$0.3\,mag above an empirical isochrone; 20\% of stars) are
%     orange.  Over $0.6<\bpmrpo<1.5$, the detected rotation periods in
%     binaries tend to fall below the slow sequence.
% 		\label{fig:binarity}
% 	}
% \end{figure}
% 
% Figure~\ref{fig:binarity} shows the rotation-color diagram for \cn\
% members, with the points colored according to indicators of binarity.
% The binaries tend to be redder and have shorter rotation periods.  We
% know of two possible explanations based on selection effects, and two
% based on physics.
% 
% Possible selection effects include {\it i)} binaries have a component
% that contributes additional red light to the system, which could skew
% the color measurement of the primary; and {\it ii)} the unresolved
% binary companions could contaminate the rotation period measurement
% \citep[{\it e.g.},][Section~5.1]{stauffer_rotation_2016}.
% 
% The two possible physical effects that could be relevant are tidal
% locking and pre-main-sequence disk locking.  Tidal locking has been
% argued to be an unlikely explanation for the frequency of rapid
% rotators due to the population statistics; a more likely scenario
% would be that the presence of the binary leads to faster disk
% dispersal, enabling the primary to contract to more rapid rotation
% periods than possible for single stars
% \citep{meibom_effect_2007,bouma_cluster_2020}.

\section{The Transit Asymmetry}
\label{app:asymmetry}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.64\textwidth]{f9.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Transit model residuals through time (binned by Kepler quarter)}.  
    {\it Left:}
    Phase-folded transit of Kepler 1627b, with stellar variability
    removed.  Black points are binned to 20 minute intervals.  The
    2-$\sigma$ model uncertainties and the maximum {\it a posteriori}
    model are shown as the faint purple band, and the dark purple
    line.
    {\it Right:}
    As on the left, with the transit removed.  Quarters 6 and 7 show a
    consistent deviation in the second half of the transit.
		\label{fig:phasequarter}
	}
\end{figure}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.64\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Transit model residuals through time (binned by year of observation)}.  
    {\it Left:}
    Phase-folded transit of Kepler 1627b, with stellar variability
    removed.  Points and models are as in
    Figure~\ref{fig:phasequarter}.
    {\it Right:}
    As on the left, with the transit removed.
		\label{fig:phaseyear}
	}
\end{figure}

\begin{figure}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Transit models and residuals, binned by quartiles in the
    local slope of the light curve}.  
    Representative uncertainties for the black points (binned at 20
    minute intervals) are shown in the lower right of each panel.  A
    similar transit asymmetry to that shown in Figure~\ref{fig:lc}
    seems to be present in three of the four bins.
    \vspace{-0.5cm}
		\label{fig:phaseslope}
	}
\end{figure}
\begin{figure}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f12.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Transit models and residuals, binned by quartiles of transit
    timing variations}.  The data are otherwise the same as in
    Figure~\ref{fig:phaseslope}.
		\label{fig:phasettv}
	}
\end{figure}

\subsection{How Robust is the Asymmetric Transit?}

As a means of exploring the robustness of the transit asymmetry,
Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear},~\ref{fig:phaseslope},
and~\ref{fig:phasettv} show the Kepler data binned in four ways: over
Kepler quarters, Julian years, quartiles of local slope, and quartiles
of transit timing variation.  Over Kepler quarters
(Figure~\ref{fig:phasequarter}), Quarter 6 shows the strongest
asymmetry: a deviation of about 3\,ppt from expectation.  Quarter 7
shows an anomaly at roughly the same transit phase.  Year 2
correspondingly shows the strongest anomaly out of any year in
Figure~\ref{fig:phaseyear}; the asymmetry is visually apparent however
in each of Years 2, 3, and 4.

To bin by quartiles in local slope and TTV, we used our measurements
of the local linear slopes in each of the observed transit windows
(144 transits total).  Four outlier transits were removed, leaving 140
transits.  These were then divided into quartiles, so that each panel
shows 35 transits binned together.  The exact light curve slope and
TTV intervals are listed in the lower left panels of
Figures~\ref{fig:phaseslope} and~\ref{fig:phasettv}.  Binned by local
slope quartiles (Figure~\ref{fig:phaseslope}), the asymmetry is
visually present in three of the four quartiles: the only bin in which
it does not appear is ${\rm d}f/{\rm d}t\in[0.001,0.033]\,{\rm
ppt\,day}^{-1}$.  The corresponding bins in TTV (largest TTV
corresponding to smallest ${\rm d}f/{\rm d}t$) show the same
asymmetries, though with somewhat greater scatter
(Figure~\ref{fig:phasettv}).

Within the theory presented by \citet{mazeh_time_2015}, unresolved
starspot crossings cause the correlation between TTVs and the local
light curve slope (Figure~\ref{fig:corr}).  In this model, we would
expect the light curves with the most negative local slopes to have
the largest positive TTVs, due to spot crossing events during the
latter half of transit.  The lower-right panel of
Figure~\ref{fig:phasettv} and upper-left panel of
Figure~\ref{fig:phaseslope} agree with this expectation.  However, we
would also expect the sign of the effect to reverse when considering
the most positive local slopes, and the most negative TTVs.  The
upper-left panel of Figure~\ref{fig:phasettv} and lower-right panel of
Figure~\ref{fig:phaseslope} contradict this expectation: the residual
in both cases maintains the same parity!  On the one hand, this shows
that the residual is not dependent on the local light curve slope,
which mitigates the possibility that it might be an artifact of our
detrending methods.  On the other, it raises the question of whether
unresolved starspot crossings are indeed the root cause of the
correlation shown in Figure~\ref{fig:corr}.

% FIXME: KM to discuss injection recovery...

\subsection{Interpretation}

The transit asymmetry seems robust against most methods of binning the
data, though with some caveats ({\it e.g.}, the ``middle quartile'' in
local flux (${\rm d}f/{\rm d}t\in[0.001,0.033]\,{\rm ppt\,day}^{-1}$),
where the asymmetry does not appear).  Nonetheless, if the effect were
entirely systematic we would expect to see it change as a function of
the sign of the local slope or TTV.  We therefore considered four
possible astrophysical interpretations: gravity darkening, transit
timing variations, spot-crossing events, and a persistent asymmetric
dusty outflow.

Gravity darkening is based on the premise that the rapidly rotating
star is oblate, and brighter near the poles than the equator
\citep[{\it e.g.},][]{masuda_spin-orbit_2015}.  The fractional transit
shape change due to gravity darkening is on the order of $(P_{\rm
break}/P_{\rm rot})^2$, for $P_{\rm break}$ the break-up rotation
period, and $P_{\rm rot}$ the rotation period.  Using the parameters
from Table~\ref{tab:posterior}, this yields an expected 0.14\%
distortion of the $\approx$1.8\,ppt transit depth: {\it i.e.}, an
absolute deviation of $\approx$2.5\,ppm.  The observed residual has a
semi-amplitude of $\approx 50\,{\rm ppm}$.  Since the expected signal
is smaller than the observed anomaly by over an order of magnitude,
gravity darkening seems to be an unlikely explanation.

The scenario of transit timing variations (TTVs) producing the
asymmetry seems unlikely because our analysis showed that transit
timing variations in \sn\ have amplitudes smaller than $\approx$10
minutes (see Figure~\ref{fig:corr}).  This is consistent with previous
results from \citet{holczer_transit_2016}.  Moreover, the transit
timing variations we do observe are correlated with the local light
curve slope, which on average increases as much as it decreases.  From
our analysis, the mean TTV and its standard deviation are
$0.00\pm12.05$ minutes; similarly the mean local slope and its
standard deviation are $0.49\pm44.49$\,ppt\,day$^{-1}$.  There is
therefore no expectation for TTVs to produce the asymmetry.  A
separate line of argument comes from Figure~\ref{fig:phaseslope}.  If
the local slope were essential to producing the transit asymmetry, we
would expect that in the largest ${\rm d}f/{\rm d}t$ bin, ${\rm
d}f/{\rm d}t\in[0.033,0.092]\,{\rm ppt\,day}^{-1}$, the sign of the
asymmetry would reverse.  We do not see evidence for this being the
case.

The third and related possibility is that of starspot crossings.
Young stars have higher spot-covering fractions than old stars ({\it
e.g.}, \citealt{morris_relationship_2020}).  Young solar-type stars
may also host dark starspots at high stellar latitudes \citep[{\it
e.g.}, EK~Dra;][]{strassmeier_starspots_2009}, though interferometric
imaging of spotted giant stars has shown different starspot latitude
distributions than those inferred from Doppler imaging
\citep{roettenbacher_contemporaneous_2017}.  Regardless, for any
spot-crossing anomalies to add coherently over the 144 Kepler
transits, it seems likely that we would need either for spots to be
persistent at a particular latitude (and for the planetary orbit to be
somewhat misaligned), or for a ``stroboscopic'' longitudinal phasing
\citep[{\it e.g.},][]{dai_stellar_2018}.  For our system, $P_{\rm
orb}/P_{\rm rot} \approx 2.76$, which means that every 4 transits and
11 stellar rotations, the planet crosses over roughly the same stellar
longitude, which might enable the necessary phasing if the spot-groups
are large and long-lived.  Unfortunately, the  S/N per Kepler transit
is $\approx8$, which renders individual spot-crossing events
unresolved.  This explanation seems marginally plausible, mainly
because the expected spot-crossing anomaly amplitudes ($\approx
100\,{\rm ppm}$) resemble the observed amplitude of the asymmetry
($\approx 50\,{\rm ppm}$).  One issue with this explanation however is
that there is no reason to expect starspot crossing events to last
exactly half the transit duration.

A persistent feature of the planet itself might therefore be needed to
explain the transit asymmetry.  An asymmetric outflow from the
planet's atmosphere could at least geometrically meet the requirements
\citep[{\it e.g.},][]{mccann_2019}.  To explain the asymmetric
transit, a small, dense component would lead the planet, and a long,
more rarefied (and variable) component would trail it.  This might
also explain the slight flux decrement visible for $\sim$1 hour
pre-ingress (Figure~\ref{fig:lc}).  The amplitude of the asymmetry is
roughly in line with theoretical expectations for dusty outflows
\citep{wang_dai_2019}, and based on the planet's size, its mass is
likely in a regime where such outflows are possible.  Out of the four
explanations discussed, this one at least theoretically seems the most
plausible.  By composition, the expectation would be that the envelope
is mostly hydrogen and helium gas, with a dust or haze component
providing the broadband opacity in the Kepler bandpass.  A natural
path toward confirming or refuting this idea would be to observe
additional transits of the planet in hydrogen absorption, metastable
helium absorption, or across a broad wavelength range in the
near-infrared.

%  \subsection{Short Cadence Data}
%  
%  \begin{figure*}[t]
%  	\begin{center}
%  		\leavevmode
%  		\includegraphics[width=1\textwidth]{f12.pdf}
%  	\end{center}
%  	\vspace{-0.7cm}
%  	\caption{
%  		{\bf Phase-folded short cadence transit of Kepler\,1627 Ab}.  
%      The precision with which the impact parameter can be measured is
%      higher than from the long cadence data.
%  		\label{fig:phase_slc}
%  	}
%  \end{figure*}
%  
%  Figure~\ref{fig:phase_slc} shows the result from analyzing 100 days of
%  short cadence (1-minute) data acquired during Quarter 15.
%  %TODO: get quantitative on better impact parameter...

\section{Spectroscopic Transit Analysis}
\label{app:rm}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=\textwidth]{f13.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Spectroscopic activity indicators during the transit of 2021
    Aug 7.} The {\it top panels} show the median line profiles Ca K,
    Ca H, and H$\alpha$ line profiles from the HIRES spectra.  The
    {\it lower panels} show the differences of each individual
    spectrum relative to the median spectrum.  The bump in the red
    wing of Ca H is H$\epsilon$.  The spectra in the lower panels are
    smoothed for visualization purposes.
    \label{fig:rvactivity}
	}
\end{figure*}

We monitored \sn\ with Keck/HIRES before, during, and after transit on
the night of 2021 Aug 7.  We used the iodine cell for wavelength
calibration\footnote{We considered a line-profile based analysis in
the regions without iodine, but the line profile stability of HIRES
precludes such an approach.}, and extracted the 1-D spectra using the
standard California Planet Survey pipeline \citep{howard_cps_2010}.
The airmass ranged between 1.1 and 2.2 from the start through the end
of observations; the seeing ranged from $1\farcs1$ at the beginning to
$1\farcs5$ at the end.  We also simultaneously observed across {\it
griz} bands using MuSCAT3 at Faulkes Telescope North on Maui, HI.

We considered two approaches to measuring the velocities: in the
first, hereafter ``Method 1'', we cross-correlated against a template
found via spectral classification with \texttt{SpecMatch-Emp}
\citep{yee_SM_2017}.  In ``Method 2'', we used a high S/N template of
V1298~Tau.  Although V1298 Tau is cooler than Kepler~1627A by
$\approx$500\,K, it has a comparable amount of line-broadening ($v\sin
i = 23$\kms), and a comparable level of stellar activity.  The mean
and standard deviation of the internal RV uncertainties averaged over
all epochs were $16.2\pm1.1$\ms\ from Method 1, and $12.6\pm0.6$\ms\
from Method 2.  The corresponding time-averaged reduced $\chi^2$ from
the template match was $1.57\pm0.04$ (Method 1) and $1.30\pm0.02$
(Method 2).  Given these diagnostics, we adopted the velocities from
the second approach.

Figure~\ref{fig:ground} shows the results.  The MuSCAT3 photometry
shows the expected starspot trend, along with the transit and what is
likely a chromatic starspot crossing event at ${\rm JD} - 2459433 =
0.955$.  The radial velocities decrease by $\approx$250\ms\ over the
six hour window.  This decrease in RV is correlated with a decrease in
the S-indices derived from the Ca HK lines.  One outlying RV point is
apparent shortly before egress; it is temporally coincident with an
outlying value in the S-index time series.

Overall, we expect the dominant trends in both the photometry and
radial velocities to be caused by starspots on the stellar photosphere
rotating into and out of view.  The plasma in the leading and receding
limbs of the stellar disk has an apparent line-of-sight velocity of
$\pm 20$\kms.  Over 10\% of a rotation cycle ($P_{\rm rot}=2.6\,{\rm
days}$), spots near these limbs come into and out of view, modulate
the stellar velocity profile, and can thereby produce the overall
$\approx$250\ms\ trend.  The Ca HK and H$\alpha$ emission profiles
support this interpretation; Figure~\ref{fig:rvactivity} shows that
each line gradually decreases in intensity over the course of the six
hour sequence.

The expectation however is for the starspot-induced signals to be
smooth, at worst with contributions at $0.5\,P_{\rm rot}$ or
$0.25\,P_{\rm rot}$ \citep{klein_simulated_2020}.  We therefore fitted
the RVs using the \citet{hirano_analytic_2010,hirano_2011} models for
the Rossiter-McLaughlin (RM) effect, and allowed for an optional
linear and quadratic trend in time to fit the $\approx$250\ms\
spot-induced trend.  We followed the methodology developed by
\citet{stefansson_2020}.  We allowed the sky-projected obliquity, the
projected stellar equatorial velocity, and the Gaussian dispersion of
the spectral lines to vary, and fixed the limb-darkening using the
$V$-band tabulation from \citet{claret_gravity_2011}.  We assumed a
Gaussian prior on $v\sin i$ and $a/R_\star$ from
Table~\ref{tab:starparams}, and also allowed for a white-noise jitter
term to be added in quadrature to the measurement uncertainties.  We
used a 15\,minute exposure time to numerically evaluate the model.

The quadratic model with the RM effect is shown in
Figure~\ref{fig:ground}; the jitter term is incorporated in the model
uncertainties, but not the plotted measurement uncertainties.  The
plotted measurement uncertainties are the internal uncertainties on
the RVs ($\approx 13\,{\rm m\,s}^{-1}$), and are dominated by the
$v\sin i$ broadening.  However, between exposures, the RVs show
significant additional scatter that is not captured by the slow
quadratic trend.  The white-noise jitter for this particular model is
$\sigma_{\rm RV} = 27^{+6}_{-5}$\ms, which is comparable to the
expected RM anomaly of $\Delta v_{\rm RM} \approx f_{\rm LD} \cdot
\delta \cdot v\sin i \cdot \sqrt{1-b^2} \approx 20\,{\rm m\,s}^{-1}$,
assuming a perfectly aligned orbit.

The presence of this additional scatter prevents a convincing
detection of the RM effect.  The reason can be understood via model
comparison.  If we compare the model with a quadratic trend and the RM
effect against a model with a linear trend and the RM effect, or even
a model with no RM effect at all, then the respective Bayesian
Information Criterion (BIC) values are as follows.
% from rm_TEMPLATE_V1298TAU_quadraticfreejitter_v0.log and
% rm_TEMPLATE_V1298TAU_v1_linearfreejitter_trendonlyfreejitter.log
\begin{align}
  {\rm BIC} &= 227.1  \quad ({\rm Quadratic+RM}) \nonumber \\
  {\rm BIC} &= 231.1  \quad ({\rm Linear+RM}) \nonumber \\
  {\rm BIC} &= 221.4  \quad ({\rm Only\ Quadratic}).
\end{align}
There is therefore no evidence to prefer the model with the RM effect
against a model that only accounts for the stellar variability.  The
``only quadratic'' model does particularly well because it can inflate
the jitter term to account for scatter during the transit (even if the
scatter contains astrophysics!), and it has fewer free parameters.
However, we cannot justify a physical prior on the jitter term,
because we do not understand the origin of the exposure-to-exposure
scatter.  As noted above, the velocity deviations from starspots are
expected to have contributions at the stellar rotation frequency, or
harmonics thereof.  This jitter is present on the exposure timescale
(15 minutes), which is only 0.4\% of the stellar rotation period; it
is not obvious that starspots would be the culprit.

The amplitude of both the spot-induced trend and the jitter are
somewhat larger than recent comparable measurements in systems such as
AU~Mic \citep{palle_transmission_2020}, DS~Tuc
\citep{montet_young_2020,zhou_well_2020} and TOI~942
\citep{wirth_2021_toi942}.  One possible explanation for the jitter is
that it is astrophysical in origin, and that it is caused by some
novel process operating on the surface of Kepler 1627A.  Another
possibility is that our RV analysis underestimates our measurement
uncertainties; in order to achieve the requisite time-sampling the S/N
per resolution element in our spectra was 70 to 80, which is lower
than desired for deriving high-precision velocities.  In addition, the
rapid rotation of the star could affect accuracy of the uncertainties
from the velocity extraction.  Observations at higher S/N are
necessary to distinguish these two possibilities, and remain
worthwhile in order to clarify the orbital geometry of \pn.  Useful
next steps would include transit observations with a stabilized
spectrograph in the optical \citep[{\it
e.g.},][]{gibson_2016_kpf,seifahrt_maroon-x_2018}, or in the
near-infrared \citep[{\it e.g.},][]{feinstein_halpha_2021}.

\section{Flare Analysis}
\label{app:flare}

In addition to the 3.9 years of long cadence data, short cadence
(1-minute) Kepler observations were acquired over 97.7 days during
Quarter 15.  The short cadence light curve shows a higher rate of
flaring than visible in the long cadence data
(Figure~\ref{fig:flarezoom}).  We analyzed the short cadence light
curve and its flares according to the following procedure.
\begin{enumerate}
  \item Fit the starspot-induced variability using a Gaussian Process
    with a \texttt{SHOTerm} kernel, a white-noise jitter term, and the
    mean flux.
  \item Select points more than twice the median absolute
    deviation from the residual, and exclude them from the light
    curve (these points include the flares).  Repeat Step~1.
  \item Using the residual from Step 2, identify all flares, requiring
    them to be at least 20 cadences apart, at least 7 median absolute
    deviations above the median baseline, and lasting at least 2
    cadences in duration.  Build the mask spanning these times, from 5
    minutes before each flare begins to 2.5 minutes after the final
    flare cadence.  Repeat Step 1 a final time.
\end{enumerate}
The final step of flare identification and fitting was performed using
\texttt{altaipony} \citep{davenport_2016,ilin_flares_2021}.  The
analytic flare model is from \citet{davenport_2014} and it
parametrizes the flare with a start time, an exponential lag time, and
an amplitude.
%Figure~\ref{fig:flarephase} shows the resulting flares, amplitudes,
%and phases.

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f14.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Flares in Kepler\,1627}.  
		{\it Top:}
		The full short cadence Kepler dataset, acquired at 1-minute
		sampling (black points) is shown with a stellar variability model
		(blue line).
		{\it Middle:}
		Residual after subtracting the stellar variability model.  Flares
		appear as spikes.
		{\it Bottom:}
		Zooms of the brightest, and third-brightest flares.  A timing
		coincidence -- that both flares have ``successors'' approximately
		one orbital period after the initial event -- is emphasized.
		\label{fig:flarezoom}
	}
\end{figure*}

%\begin{figure*}[t]
%	\begin{center}
%		\leavevmode
%		\includegraphics[width=0.9\textwidth]{f8.pdf}
%	\end{center}
%	\vspace{-0.7cm}
%	\caption{
%		{\bf Phase-folded flares in Kepler\,1627}.  
%    {\it Top:}
%    As in the middle of Figure~\ref{fig:flarezoom}, phase-folded at
%    the planet's ephemeris.
%    {\it Bottom:}
%    Fitted flare amplitudes and orbital phases -- each point
%    represents one flare.
%    Colors indicate relative time, from the beginning of the
%    98-day short cadence Quarter 15 dataset (dark blue) to the end (light
%    yellow).
%    Lower amplitude flares likely exist in the data, and were not
%    examined.
%		\label{fig:flarephase}
%	}
%\end{figure*}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f15.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Statistics of inter-flare arrival times}.  
    24 flares were recorded with amplitudes exceeding 0.5\% over the
    97.7 days of short cadence observations.  The histogram of the
    time intervals between every possible pair of flares is shown in
    black.  Some plausibly important timescales for star-planet
    interactions, namely the planetary orbital period and synodic
    period (the orbital period as seen from the rotating stellar
    frame) are shown along with their linear combinations.  Monte
    Carlo draws from a Poisson distribution are shown with the gray
    bands.  While peaks in the observed distribution do coincide with
    some of the ``special periods'', the statistical evidence for a
    non-Poissonian process driving the flares does not reach the
    $5$-$\sigma$ threshold.
		\label{fig:flarestats}
	}
\end{figure*}

There were $N_{\rm f}=24$ flares that exceeded $0.5\%$ in relative
flux during the short cadence observations.  These 24 flares spanned a
total of 6.5 hours ($\sim$15 minutes per flare).  Inspecting the data,
we noticed a coincidence in the flare arrival times.  The coincidence
is that despite the low flare duty cycle, one orbital period after the
brightest flare, a second flare followed.  This and a similar event
are shown in Figure~\ref{fig:flarezoom}.  The timing error is good to
a $\approx0.2\%$ difference from the orbital period, which given the
duty cycle seems {\it a priori} unlikely.  If we consider flares
falling within 2\% of the planet's orbital period after a previous
flare, then 4 of the 24 flare events have candidate ``successors''.

As with any coincidence, if one does not have a firm prediction, it is
difficult to assess whether a surprise is statistically significant.
Since our surprise was specifically at the inter-arrival time of
certain flares coinciding with special time intervals, we performed
the following analysis.  First, we considered all unordered pairs of
flares.  For $N$ flares there are ${n \choose 2}$ such pairs (for our
case, 276 pairs).  We then compared the distribution of the pair
separations against that of a Poisson distribution.  Specifically, we
drew $N_{\rm f}=24$ samples from a Poisson distribution with $\lambda
= \Delta t / N_{\rm f}$, for $\Delta t=97.7\,{\rm days}$ the total
duration of the observations, and repeated the draw $10^3$ times with
unique random seeds.

Figure~\ref{fig:flarestats} shows the results.  The vertical lines in
the figure show the planetary orbital period, the synodic period
$P_{\rm syn} = (P_{\rm rot}^{-1} - P_{\rm orb}^{-1})^{-1}$, and linear
combinations thereof.  The tidal period (half the synodic period) is
not shown.  The bins are logarithmically spaced to give 100 bins
between the minimum and maximum ordinate values.  The gray bands
express the range of values observed from the Poissonian draws.  While
it does seem like an odd coincidence for peaks in the observed flare
arrival time distribution to coincide with the locations of these
``special intervals'', the statistical evidence for a non-Poissonian
process driving the flares does not seem especially overwhelming.
More quantitatively, the peaks observed at the orbital and synodic
periods are within the $\pm 2$-$\sigma$ range of a Poissonian process,
and those at $P_{\rm orb}+P_{\rm syn}$ and $P_{\rm orb}+2P_{\rm syn}$
are only slightly above this range.  With that said, future analyses
of these data by investigators with more knowledge of this topic could
very well yield more quantitative insights.  Such analyses should keep
in mind an important caveat: the amplitude distribution of M-dwarf
flares extends up to many times the quiescent flux \citep[see Figure~7
of][]{gunther_2020}.  A flare on Kepler\,1627B producing double its
quiescent white-light flux would yield a $\approx$1\% apparent
amplitude.  Such flares could represent a significant fraction of
those in the Kepler observations.

\listofchanges

%\allauthors
\end{document}