### 2021-11-24 paper_review/response_1.txt
stage paper_review

Dear Dr. Jackson,

We thank you for organizing the review of our paper, and also thank
the referee for constructive criticism.  We have considered the
referee's remarks carefully and revised our manuscript accordingly.

Below, we have reproduced the relevant portions of the referee's
report, alongside our responses. A list of changes is appended to
the manuscript using the trackchanges AASTeX macros.

Sincerely,

Luke Bouma

******************************************
Reviewer #1:

REVIEWER COMMENT
----------
Line 57 (evaporation): Consider adding citation to Lopez et al. (2012) or
similar

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Line 78 (Kepler ages): Consider adding citation to Walkowicz & Basri (2013)

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Line 93: Clarify origin of name delta Lyr, given there are two pre-existing
names, and how they overlap. It might also be worth mentioning Kerr's Cep-Her
association as in prep just to save future readers the headache of
cross-identifying them, either here or in Sec 2.1

RESPONSE
----------
> delta Lyr cluster was the first name suggested by Stephenson (1958),
> and it is the default name listed in SIMBAD.  It was also the name
> used in the title of Eggen's 1967 paper.
>
> The cluster was called Stephenson-1 by people other than Stephenson,
> including by Kounkel & Covey 2019 -- their Theia catalog identifiers
> are simply given to any cluster in their catalog, in addition to any 
> previous names that existed.
>
> The overlap between the earlier lists (Stephenson 1958, Eggen 1968,
> Kharchenko et al. 2012) and the Theia 73 list from Kounkel & Covey
> 2019 is only present in the stars closest to delta Lyr itself.  For
> instance, see the 30 stars shown in Figure 1 of Eggen 1968.  The
> broader "complex" shown in Figure 1 of our paper, which to our
> knowledge first reported to be a set of kinematically associated
> stars by Kounkel & Covey (inasmuch as an "HDBScan group"
> implies kinematic association).
>
> Regarding Ronan Kerr's view of the delta Lyr cluster as a component
> of the larger Cep-Her complex, we would prefer to leave this as
> Ronan's result to present!  There's significant work here in
> designing the clustering method and verifying that it works.
> There's also work in understanding the overall structure of the
> broader association (which is different in nuanced ways from the
> gray Theia 73 points in Figure 1).  Given the complexity, and the
> potential importance of the complex given its proximity and age, we
> really would prefer to leave this for the in prep R. Kerr work.  And
> similarly, the introduction of the term "Cep-Her complex" is
> probably best left for that future work.

REVIEWER COMMENT
----------
Line 165: Consider mentioning the axes on which data was cleaned (e.g. color
precision)

RESPONSE
----------
> Done - thank you.  We have added the following sentence: "This also
> involved cuts on the photometric signal to noise ratio, the number
> of visibility periods used, the astrometric $\chi^2$ of the
> single-source solution, and the $G_{\rm BP}-G_{\rm RP}$ color excess
> factor."

REVIEWER COMMENT
----------
Line 179 (reddening): What does "reasonably well" mean? Something like 1-2
sigma?

RESPONSE
----------
> Thank you for catching the weasel words.  The revised text now reads
> as follows: "These values \replaced{agree reasonably well with}{are
> within a factor of two of} previously reported values
> \replaced{from}{in} the literature \citep[{\it
> e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}\added{,
> and are all small enough that the choice of whether to use them {\it
> vs.} other extinction estimates does not affect our primary
> conclusions}."
>
> More quantitatively, from the Lallement maps we report
> $E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$ for IC
> 2602, the Pleiades, and the delta Lyr cluster.  The relevant
> literature reports:
>
>    Gaia+2018 E(B-V):
>    Pleiades 0.045
>    IC2602 0.031
>   
>    Randich+2018 E(B-V):
>    IC2602 0.032 - 0.092
>   
>    KC2019: E(B-V)
>    Pleiades 0.071
>    IC2602 0.078
>    del Lyr cluster 0.059
>   
>    Bossini+2019 E(B-V)
>    Pleiades/Melotte22 0.045
>    IC2602 0.031
>    del Lyr cluster 0.051
>
> hence our assessment of "it's within a factor of two, and doesn't
> matter much."  Given that all these clusters are within ~400
> parsecs, this perhaps isn't too surprising.

REVIEWER COMMENT
----------
Fig 3: It may be helpful to include one of the even younger clusters on the CMD
(i.e. to bracket delta Lyra), though perhaps it makes the plot too busy.

RESPONSE
----------
> Please see camd_with_UCL.png and camd_with_UCL_no_redcorr.png
> attached.  This approach indeed makes the plot a bit too busy, in
> our opinion.

REVIEWER COMMENT
----------
Page 3: Consider re-organizing. Based on the 1st paragraph, I thought the
quantitative analysis would only deal with Pleiades and IC 2602 and one might
infer reddening was only calculated for these clusters (whereas I assume the
latter are also corrected for reddening prior to the analysis described). My
suggestion would be to introduce all the clusters first. It would also be
helpful to list the age of IC 2602 at some point in the text.

RESPONSE
----------
> Regarding the main re-organization point, FIXME TODO REORGANIZE?
>
> Regarding the last point, the final paragraph of Section 2.2.1
> discusses the age of IC 2602 in the text.  There's a fairly wide
> spread of reported literature ages between 30 and 46 Myr -- we
> ultimately adopted the half-way point between the isochrone and LDB
> ages.

REVIEWER COMMENT
----------
Line 202 (binaries): Add RUWE citation(s). Nice approach to remove
radial_velocity_error based on each cluster's distribution. It should be noted
somewhere in this discussion that the target itself does *not* meet the RUWE
limit and so wouldn't appear on the plot by default.

RESPONSE
----------
> We've added a footnote to the Gaia DPAC memo by Lindegren on the
> RUWE.
>
> The stars shown in Figure 3 are those discussed in the first
> paragraph of Section 2.2.1.  (The binarity cut has not yet been
> applied -- which is why the binary sequence is visible for the δ Lyr
> cluster).  So for the latter point, we have not made any changes.

REVIEWER COMMENT
----------
Line 209: I don't follow starting "assembled a piecewise grid of hybrid
isochrones..." What does "hybrid" mean here? Once the splines for each cluster
are defined, how does one interpolate between them? Are the ages then
determined for each star and combined?

RESPONSE
----------
> The manuscript relies on the citation to Gagne+2020 for the full
> description of this method.  The "hybrid" isochrone I_i is the
> linear combination of any two reference isochrones.  The cited
> Equation 6 of Gagne+2020's mu-Tau paper gives the definition:
>
>   I_i = α_i * I_45 + (1-α_i)*I_112,
>
> for α_i a coefficient between 0 and 1, I_45 the reference 45 Myr
> isochrone spline (from Tuc-Hr, Columba, and Carina in Gagne's
> paper), and I_112 the Pleiades isochrone spline. 
>
> In our case, given our different reference clusters, the definition
> is modified to
>
>   I_i = α_i * I_16 + (1-α_i)*I_38,   (age from 16-38 Myr)
>   I_i = α_i * I_38 + (1-α_i)*I_112,  (age from 38-112 Myr)
>
> where I_38 is the IC 2602 isochrone spline, and I_16 is the UCL
> isochrone spline.
>
> The main assumption of this approach is that between the reference
> times set by the adopted clusters, the isochrones evolve linearly in
> time through color-absolute magnitude space.  If we were to attempt
> this linear interpolation all the way from UCL to the Pleiades
> without putting any intermediate point, the validity of this
> assumption would be crucial to test.  Since we include IC 2602,
> and it overlaps almost directly on the CAMD with the δ Lyr cluster,
> the only useful quantity that this method really gives us is a
> formal uncertainty on the age.  (Since the age the method returns
> will be very close to whatever age we adopt for IC 2602).
>
> To better illustrate the method, we've attached a figure to the
> response, "age_interpolation_example.png". This shows the CAMD data
> from the clusters after all the cleaning steps, their reference
> splines (constructed from bins over a select color range), and the
> interpolated splines spaced every 4.4 Myr between UCL and IC2602,
> and every 14.8 Myr betwen IC2602 and Pleiades.  While we do interpolate
> over a much finer grid, this is not shown to improve the clarity of
> the figure.  The age of δ Lyr is determined by using all the stars
> in the displayed color range, through the referenced Equation 7 of
> Gange+2020, which is a Gaussian likelihood that treats the
> interpolated isochrone as the "model" and the δ Lyr cluster's
> isochrone as the "data".  The statistical uncertainty on the age is
> then the standard deviation of a Gaussian fit to the posterior
> probability (in log-age; it becomes asymmetric when transformed to
> linear age).  We've added a few sentences to the manuscript to
> clarify this latter point.

REVIEWER COMMENT
----------
Line 230 (age): Consider adopting a systematic error (perhaps corresponding to
the range noted for the extremes of the ages) in addition to the random
(assumed to be the quoted +6/-5 error). Relatedly, how are the quoted errors
derived?

RESPONSE
----------
> We're comfortable quoting the statistical error derived from the
> standard deviation of the logarithmic age posterior since this
> distribution is roughly Gaussian.  The worry with quoting a
> systematic error, e.g., "The age of the δ Lyr cluster is
> $38^{+6}_{-5}$ (stat) $^{+6}_{-8}$ (sys)" is that the systematic
> uncertainties are not Gaussian and are not based on any probability
> distribution that we can easily assess -- they're estimated based on
> the extremes of previous literature age measurements.  Quoting this
> systematic range next to a number that has been derived based on a
> more formal approach (and that has statistical meaning within that
> approach) somehow misses the point.  Given the roughly comparable
> amplitude of the two sources of uncertainty, does it matter?  Open
> cluster ages are not H_0.

REVIEWER COMMENT
----------
Line 235: May be worth mentioning the restriction to the 391 stars in the
beginning of sec 2.2.2.

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Line 278: "Indistinguishable" -- from each other or from the Pleiades?

RESPONSE
----------
> From each other (and the manuscript has been updated; thank you!).

REVIEWER COMMENT
----------
Line 286: I thought I remember Curtis or Douglas citing binarity as a
possibility for stars above to slow sequence as well, but I don't have a
reference on-hand.

RESPONSE
----------
> Thanks for this point -- we've updated the text with a reference to
> Stauffer+16.  In their Section 5.1, they discuss the case of some
> unresolved F+K binaries in the Pleiades which show anomalously long
> K2 rotation periods.

REVIEWER COMMENT
----------
Section 3.2: The text mentions statistical and systematic uncertainties but I
only see one set of errors listed in Table 1. Were these uncertainties combined
or was only one adopted in the end?

RESPONSE
----------
> They were combined in quadrature (see Line 331 of the original
> manuscript, "Reported uncertainties are a quadrature sum of the
> statistical and systematic components."

REVIEWER COMMENT
----------
Line 391: It would be helpful to state the star's RUWE in the text here.
Additionally, I believe RUWE acornym was already defined previously.

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Section 3.2: Comment on whether the binary companion should effect the rotation
measurement of the host star (either here, or in the rotation section).

RESPONSE
----------
> We've added the following sentence at the beginning of Section 4.1,
> when the Kepler light curve is introduced
>
> """
>The resulting photometry is dominated by a quasi-periodic starspot
>signal with a peak-to-peak amplitude that varies between 2\% and 8\%.
>\added{Given that the secondary companion's brightness in the Kepler
>band is 1\% to 2\% that of the primary, source confusion for the
>rotation signal is not expected to be an issue.}
 """
> The rationale here is that the rotation amplitude on the secondary
> would need to be a factor of >two times its intrinsic brightness for
> any confusion to happen, and the primary would need to not show any
> rotation signal.  Such large variability amplitudes have not been
> observed except for disked stars (e.g., Rebull+2020, Figure 10), and
> there is no evidence for a secondary period in the Lomb Scargle
> periodograms.

REVIEWER COMMENT
----------
One note about the transit analysis: The MuSCAT3 data was introduced as
supporting evidence but the conclusion (that the transits appear achromatic)
was never explicitly stated.

RESPONSE
----------
> Line 642 of the original manuscript notes "Although we did not
> detect the RM anomaly, the multi-band MuSCAT3 light curves show that
> the transit is achromatic (Figure 7)."

REVIEWER COMMENT
----------
Line 725: The fast rotation period and flares could plausibly effect the
completeness.

RESPONSE
----------
> These completeness statistics were computed using KeplerPORTS (the
> Burke & Catanzarite 2021 citation).  KeplerPORTS does correct for
> the red noise in the light curves by using the CDPP over different
> timescales -- see Section 3.3 of Burke, C. J. & Catanzarite, J.
> 2017, Planet Detection Metrics: Per-Target Detection Contours for
> Data Release 25, KSCI-19111-002.  Nonetheless, we do agree that this
> is a worry -- perhaps a full injection-recovery analysis of the
> light curve with a pipeline of our own might help verify whether
> these detection limits are plausible.  This effort is outside our
> scope, for now.

REVIEWER COMMENT
----------
Line 729: Considering Owen 2020, contraction could also contribute to some
extent to the larger radii of small planets; might this be significant?

RESPONSE
----------
> TODO WHAT?

REVIEWER COMMENT
----------
Appendix A paragraph beginning at line 1158: May be worth mentioning that delta
Lyr may also be part of one of these heirarchical associations.

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Figure 13 caption: Specify planetary orbital period since technically there is
also a stellar orbital period involved.

RESPONSE
----------
> Done, thank you.

******************************************
Reviewer #2:

REVIEWER COMMENT
----------
Abstract
line 24: Kepl-1627Ab cannot be a "mini-Neptune" if it is same size of Neptune.
Replace with "Neptune-size" Yes, it falls within a distribution of objects in a
period-radius diagram with an average that is smaller than Neptune, but
claiming affiliation with smaller objects implies more than we know about this
planet, e.g. its mass.

RESPONSE
----------
> Point well-taken. 
> R_p = 3.79 +/- 0.16 R_earth, so yes, this is the same size as
> Neptune within 1-sigma.

REVIEWER COMMENT
----------
line 25: Kepler target stars were chosen for their suitability for transiting
detection of planets in general, and for planets that are Earth-analogs in
paraticular. That involves radius, brightess. Distance and spectral types were
not direct selection criteria; this was before Gaia.

RESPONSE
----------
>

REVIEWER COMMENT
----------
line 28: "main" --> "prime"

RESPONSE
----------
>

REVIEWER COMMENT
----------
Intro

Line 47: (First paragraph on p 2) the time-scale of 1e8 is associated with one
particular set of photoevaporation models; core-powered mass loss proposes
timescales of Gyr(s) (see the Ginzburg paper that was cited). More recent
photoevaporation models also propose 1 Gyr timescales. It would be good to
revise/update this paragraph.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 2

Line 65: it is convention to italicize names of space missions like TESS and K2

RESPONSE
----------
>

REVIEWER COMMENT
----------
Lines 142-144: perhaps rephrasing this would clarify that the authors did not
use Kepler-1627 as a fiducial star about which to select cluster members, and
that they restricted the analysis to one of several sub-clusters or clumps
which contained Kepler-1627.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Figure 1: the first three panels would be more effective if they were plotted
at the same scale and with appropriately aligned axes. As such the upper right
panels are too small to be useful. Consider leaving out the Sun (which is
irrelevant) so as to have a more appropriate choice of scale to how the details
of the cluster. The purpose of the fourth panel (tangential velocities) and the
second line in the caption are unclear to me. Aren't the tangential velocities
known from Gaia proper motions and parallaxes? Why calculate them with the
assumption of common space motion? If that is done, what does it show, anyway?

RESPONSE
----------
>

REVIEWER COMMENT
----------
Figure 3 bottom: Can the IC 2602 and 2391 rotation periods be added to this
plot? Lines 276-278 mention that these and cite Douglas et al. 2021 (although
the reference does not seem to be complete). Isn't it interesting that the
interlopers all fall on the Praesepe rotation sequence?

RESPONSE
----------
>

REVIEWER COMMENT
----------
p 5 lines 258-261: rather than simply throw the TESS values out in favor of the
Kepler ones, the authors should compare them to evaluate the accuracy and
reliability of the TESS values.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 3

page 7 line 391: report the RUWE value (yes its in the Table but don't make the
reader work)

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 4.1.3 and Figure 6

The significance of this slope is unclear and a sigma is not a good way to
evaluate it. Perform a ratio of chi-squared or F-test to evaluate it vs. a flat
line.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Lines 530-542: does option (ii) include a transit fit as well? I assume it does
but it would be good to clarify. And is option (i) the 2nd and 4th order
polynomial? Perhaps just state this?

RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 561-562: do the authors mean "lack of statistical significance"?

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 4.2

Analysis of the motion of the photocentroid has been widely used to help
validate Kepler planets. Can this be used here to distinguish between the
possible A and B host stars?

RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 604-608: The statement that the density from the transit duration is
inconsistent with that expected for the M dwarf "B" cites Table 2 and
Choi+2016, but that Table does not appear to provide any information on "B",
and the Choi+2016 reference is just for the Mesa models. A density of 2 g/cm3
(or lower) is possible for an M dwarf on the pre-main sequence, depending on
the mass, age and model used.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Lines 613-614: it would be good to quantify this, i.e. how faint an unresolved
companion or background star must be based on the star's position in the CMD.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Lines 635-638: Here, new observations appear for the first time but are not
described. While the details are given in the appendix, and the outcome of the
observations doesn't define this work, the authors need to give some basic
description of the observations, e.g. passbands, SNR, resolution, observation
times, etc. The authors should not assume the reader actually know what HIRES
and MuSCAT are!

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 5

line 670: assuming that SNR is not to blame, the problem is not precision, the
problem is either measuremnt systematics or astrophysical noise.

RESPONSE
----------
>

REVIEWER COMMENT
----------
line 678-683: The Ca II HK activity indicator "S" as well as the H-alpha
equivalent width could be obtained from the HIRES spectrum. Have the authors
checked fora X-ray or UV (Galex) detection of this star? The star is likely not
detected but its worth a look. In the same vein, the authors may wish to
comment on whether the star's infrared emission indicates or rule out a
substantial debris disk -- the star is detected in WISE (and perhaps Spitzer?)

RESPONSE
----------
>

REVIEWER COMMENT
----------
lines 699-701 It could be mentioned here that the planet orbital period and
stellar rotational period are not a rational combination.

RESPONSE
----------
>

REVIEWER COMMENT
----------
line 723-726: The caveat that should be mentioned here is that this is for
older, main sequence stars.

RESPONSE
----------
>

******************************************

Other minor changes made to the manuscript include:

* Updated binning interval on lower panels of Figure 5 from 20 to 15 minutes.

* L.G.B.'s affiliation order was reversed, to reflect the primary
  institution at which this work was performed.