### 2022-04-11 papers/ms.tex
update ms

%\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
\documentclass[12pt,twocolumn]{aastex63}
%\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

\newcommand{\cn}{Cep-Her complex} % cluster name
\newcommand{\sysone}{Kepler-1627} % star system name (binary)
\newcommand{\stone}{Kepler-1627 A} % star system name (binary)
\newcommand{\plone}{Kepler-1627 Ab} % planet name
\newcommand{\systwo}{Kepler-1643} % star system name (binary)
\newcommand{\sttwo}{Kepler-1643} % star system name (binary)
\newcommand{\pltwo}{Kepler-1643 b} % planet name
\newcommand{\systhree}{KOI-7368} % star system name (binary)
\newcommand{\stthree}{KOI-7368} % star system name (binary)
\newcommand{\plthree}{KOI-7368 b} % planet name
\newcommand{\sysfour}{KOI-7913 } % star system name (binary)
\newcommand{\stfour}{KOI-7913 A} % star system name (binary)
\newcommand{\plfour}{KOI-7913 Ab} % planet name

\newcommand{\clusterage}{$38^{+6}_{-5}$\,Myr} %

\newcommand{\npms}{1097} % 20220311_Kerr_SPYGLASS205_Members_All.csv

%
% Symbols
%
\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\mkms}{{\rm \,km\,s^{-1}}}  % math mode
\newcommand{\ms}{\,m\,s$^{-1}$}
\newcommand{\bpmrpo}{(G_{\rm BP}-G_{\rm RP})_0}
\newcommand{\bpmrp}{G_{\rm BP}-G_{\rm RP}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{\today}
\revised{---}
\accepted{---}
%\submitjournal{AAS Journals}
\shorttitle{Trio of Mini-Neptunes in Cep-Her}

\begin{document}

\title{
  %Three Mini-Neptunes in a 40 Million Year Old Behemoth
  Kepler and the Behemoth: Three Mini-Neptunes in a 40 Million Year Old Association
  %Three Mini-Neptunes, 40 to 50 Million Years Old 
  %A Trio of 40 to 50 Million Year Old Mini-Neptunes
}

%\suppressAffiliations
%\NewPageAfterKeywords
\input{authors.tex}

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a refined
  view of open cluster dispersal.  Here we present an analysis of a
  group stars spanning Cepheus ($l=100^\circ$) to Hercules
  ($l=40^\circ$), hereafter the Cep-Her complex.
  The group includes four Kepler Objects of Interest:
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), 
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$), and
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
  The latter Neptune-sized planet is in a component of the Cep-Her
  complex called the $\delta$\ Lyr\ cluster
  \citep{bouma_kep1627_2022}.  Here we focus on the former three
  systems, which are in other regions of the association.  Based on
  kinematic evidence from Gaia, stellar rotation periods from TESS,
  and spectroscopy, these three systems are also $\approx$40 million
  years (Myr) old.  More specifically, we find that Kepler-1643 is
  $46^{+9}_{-7}$\,Myr old, based on its membership in a dense
  sub-cluster of the complex called RSG-5.  KOI-7368 and KOI-7913 are
  $36^{+10}_{-8}$\,Myr old, and are in a diffuse region that we call
  CH-2.  Based on the transit shapes and high resolution imaging, all
  three objects are most likely planets, with false positive
  probabilities of $6\times10^{-9}$, $4\times10^{-3}$, and
  $1\times10^{-4}$ for Kepler-1643, KOI-7368, and KOI-7913
  respectively.  These planets empirically demonstrate that
  mini-Neptunes with sizes of $\approx$2 Earth radii exist at ages of
  $\approx$40 million years.
\end{abstract}

\keywords{
  exoplanet evolution (491),
  open star clusters (1160),
	stellar ages (1581)
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% * Main text <3500 words (not including acknowledgements, appendices, or other
%   supplementary)

\section{Introduction}

The discovery and characterization of transiting planets younger than
a billion years is a major frontier in current exoplanet research.
The reason is that the properties of young planets provide benchmarks
for studies of planetary evolution.  For instance, there are the
questions of when hot Jupiters arrive on their close-in orbits
\citep{dawson_johnson_2018}, how the sizes of planets with massive
gaseous envelopes evolve \citep{rizzuto_tess_2020}, when and if
close-in multiplanet systems fall out of resonance
\citep{arevalo_stability_2022,goldberg_architectures_2022}, and
whether and how mass-loss explains the radius valley
\citep{lopez_how_2012,Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018,lee_primordial_2021}.

The discovery of a young planet requires two claims to be fulfilled:
the planet must exist, and its age must be secured.  Spaced-based
photometry from K2 and TESS has yielded a number of young planets for
which the planetary evidence comes from transits, and the age is based
on either cluster membership
\citep{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
or else on correlates of youth such as stellar rotation, photospheric
lithium content, x-ray activity, or emission line strength
\citep{zhou_2021_tois,hedges_toi-2076_2021}.

In this work, we leverage recent analyses of the Gaia data, which have
greatly expanded our knowledge of stellar groups \citep[{\it
e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To date these
analyses have mostly clustered on the stellar positions and 2-D
velocities measured by Gaia.  One important result is the discovery of
diffuse streams and tidal tails comparable in stellar mass to the
previously known cores of nearby open clusters
\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  Even
though these streams can be spread out over tens to hundreds of
parsecs, their velocity dispersions can remain coherent at the
$\sim$1\kms\ level.  They can also be much larger: as an extreme
example, internal dynamics and projection effects in the Hyades drive
stars in the tidal tails span up to $\pm 40\mkms$ in relative velocity
when projected against the cluster center \citep{jerabkova_800_2021}.
The stars in such diffuse regions can be verified to be the same age
as the core members ({\it i.e.}, coeval) through analyses of
color-absolute magnitude diagrams \citep{KounkelCovey2019}, stellar
rotation periods \citep{curtis_tess_2019,bouma_2021_ngc2516}, and
chemical abundances \citep{hawkins_2020}.  While there are many
implications for our understanding of star formation and cluster
evolution \citep{dinnbier_tidal_2020}, a separate consequence is that
we now know the ages of many more stars, including previously known
planet hosts.

The prime Kepler mission \citep{borucki_kepler_2010} found most of the
currently known transiting exoplanets, and it was conducted before
Gaia.  It is therefore sensible to revisit the Kepler field, given our
new knowlege of the stellar ages.

In this work, we expand on our previous study of a $38^{+7}_{-6}$ Myr
old Neptune-sized planet in the Kepler field (Kepler-1627~Ab;
\citealt{bouma_kep1627_2022}).  The age of this planet was derived
based on its host star's membership in the $\delta$\ Lyr\ cluster.
Our analysis of the cluster focused on the immediate spatial and
kinematic vicinity of Kepler-1627~A in order to secure the planet's
age.  However it became clear that the $\delta$\ Lyr\ cluster seems to
also be part of a much larger group of similarly aged stars.  This
group, which is at a distance of $\sim$330\,pc from the Sun, appears
to span Cepheus to Hercules (galactic longitudes, $l$, between
40$^\circ$ and 100$^\circ$), at galactic latitudes roughly between
0$^\circ$ and 20$^\circ$.  We therefore refer to it as the Cep-Her
complex.  It exhibits significant sub-structure over its $\approx$250
parsec length, and a detailed analysis of its memberships, kinematics,
and possible origin is currently being prepared by R.~Kerr and
collaborators.

Here, our focus is on the intersection of the Cep-Her complex with the
Kepler field.  Cross-matching the full set of candidate Cep-Her
members against known Kepler Objects of Interest (KOIs)
\citep{thompson_planetary_2018} yielded four candidate cluster
members: Kepler-1627, Kepler-1643, KOI-7368, and KOI-7913.  Given our
previous analysis of Kepler-1627, we will mostly focus on the latter
three.  After analyzing the relevant properties of Cep-Her
(Section~\ref{sec:cluster}), we derive the stellar properties
(Section~\ref{sec:stars}) and validate the planetary nature of each
system using a combination of the Kepler photometry and
high-resolution imaging (Section~\ref{sec:planets}).  We conclude with
a few caveats, and discuss implications for the size-evolution of
close-in mini-Neptunes (Section~\ref{sec:disc_conc}).

\section{The Cep-Her Complex}
\label{sec:cluster}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
  {\bf Positions and velocities of candidate members of the Cep-Her
  complex.}
  {\it Top row}: On-sky positions in galactic coordinates.  Black
  points are stars for which group membership is more secure than for
  gray points.  Kepler-1627 is in the outskirts of the $\delta$ Lyr
  cluster \citep{bouma_kep1627_2022}, which is centered at $\{ l, b\}
  \approx \{ 66^\circ, 12^\circ\}$.
  {\it Middle row}: Galactic positions.  The Sun is at $\{X, Y, Z\} =
  \{0, 0, 20.8\}$\,pc; lines of constant heliocentric distance are
  shown between 250 and 400\,pc, spaced by 50\,pc.
  {\it Bottom row}: Galactic tangential velocities (left) and
  galactic longitudinal velocity versus galactic longitude (right).
  The gray band in the lower-right shows the $\pm$1-$\sigma$
  projection of the Solar velocity with respect to the local standard
  of rest.  There is a strong spatial and kinematic overlap between
  Kepler-1643 and RSG-5 (magenta).  The local population
  of candidate young stars around KOI-7368 and KOI-7913 is more
  diffuse -- we call this region ``CH-2'' (lime-green).
  The selection method for the black, gray, purple, and green points
  is described in Section~\ref{subsec:members}.
	\label{fig:XYZvtang}
	}
\end{figure*}

\subsection{Previous Related Work}

Our focus is on a region of the Galaxy approximately 200 to 500\,pc
from the Sun, above the galactic plane, and spanning galactic
longitudes of roughly $40^\circ$ to $100^\circ$ degrees.  Two rich
clusters in this region are the $\delta$~Lyra cluster
\citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
Each of these clusters was known before Gaia.  They have reported ages
between 30 and 60\,Myr.  Early empirical
evidence that these two clusters could be part of a large and more
diffuse population was apparent in the Gaia-based photometric analysis
of pre-main-sequence stars by \citet[][see their Figures~11
and~13]{Zari2018}.  Further kinematic connections and complexity were
highlighted by \citet{KounkelCovey2019}, who included these previously
known groups in the larger structures dubbed ``Theia~73'' and
``Theia~96''\footnote{See their visualization online at
\url{http://mkounkel.com/mw3d/mw2d.html} (accessed 15 March 2022)}.
The connection made by \citet{KounkelCovey2019} between the previously
known open clusters and the other groups in the region was made as
part of an unsupervised clustering analysis of the Gaia DR2 positions
and tangential velocities with a subsequent manual ``stitching'' step.
Their results support the idea that there is an overdensity of
30 to 60\,Myr old stars in this region of
the Galaxy.  \citet{Kerr2021}, in a volume-limited analysis of the
Gaia DR2 point-source catalog out to one third of a kiloparsec,
identified three of the nearest sub-populations, dubbed
``Cepheus-Cygnus'', ``Lyra'', and ``Cerberus''.  \citet{Kerr2021}
reported ages for each of these subgroups between 30 and 35 Myr.

\subsection{Member Selection}
\label{subsec:members}

The possibility that the $\delta$~Lyr cluster, RSG-5, and the
sub-populations identified by \citet{Kerr2021} share a common origin
has yet to be fully substantiated, and is the subject of the upcoming
study by R.~Kerr and collaborators.  Our primary interest in the
region stems from the fact that a portion of it was observed by Kepler
(Figure~\ref{fig:XYZvtang}, top panel).  To further explore the
population of stars that were observed, we select candidate Cep-Her
members through four steps, the first three being identical to those
described in Section~3 of \citet{Kerr2021}.  We briefly summarize them
here.

The first step is to select stars that are photometrically distinct
from the field star population based on Gaia EDR3 magnitudes $\{G,
G_{\rm RP}, G_{\rm BP}\}$, parallaxes and auxiliary reddening
estimates \citep{lallement_gaia-2mass_2019}.  This step yielded \npms\
stars with high-quality photometry and astrometry.  These stars are
either pre-main-sequence K and M dwarfs due to their long contraction
timescales, or massive stars near the zero-age main sequence due to
their rapid evolutionary timescales.

The second step is to perform an unsupervised HDBScan clustering on
the photometrically selected population
\citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The
parameters we use in the clustering are $\{ X, Y, Z, c v_b, c v_{l^*}
\} $, where $c$ is the size-velocity corrective factor, which is taken
as $c=6\,{\rm pc / km\,s}^{-1}$ to ensure that the spatial and
velocity scales have identical standard deviations.  Positions are
computed assuming the \texttt{astropy v4.0} coordinate standard
\citep{astropy_2018}, which places the Sun $8122$ pc from the galactic
center, and assumes the solar velocity with respect to the local
standard of rest from \citet{schonrich_local_2010}.  As input
parameters to HDBScan, we set the minimum $\epsilon$ threshold past
which clusters cannot be fragmented as $25$\,pc in physical space, and
$c$\,\kms\ in velocity.  The minimum cluster size $N$ is set to 10, as
is $k$, the parameter used to define the ``core distance'' density
metric.

This unsupervised clustering in our case yielded 8 distinct groups.
These groups are then used as the ``seed'' populations for the third
step, which is to search for objects at least as close to the 10$^{\rm
th}$ nearest HDBSCAN-identified member in space-velocity coordinates.
This third step yields stars that are spatially and kinematically
close to the photometrically young stars, but which cannot be
identified as young based on their positions in the color-absolute
magnitude diagram.

The outcome of the analysis up to the point of the third step is shown
in Figure~\ref{fig:XYZvtang}.  To enable a selection cut that filters
out field-star contaminants, we also compute a weight metric, defined
such that the group member with the smallest core distance has a
weight of 1, the group member with the greatest core distance has a
weight of 0, and weights for the other group members are log-normally
distributed between these two extremes.  In Figure~\ref{fig:XYZvtang},
we show 12{,}436 objects with weight exceeding 0.02 as gray points,
and overplot 4{,}763 objects with weights exceeding 0.10 as black
points.\footnote{These counts only include objects with reliable
astrometry and photometry: $\varpi/\sigma_\varpi>5$;
$G/\sigma_{G}>50$; $G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm
BP}/\sigma_{G_{BP}}>20$.} The previously known $\delta$~Lyr\ cluster
is visible at $(l,b)=(68^\circ,15^\circ$) and $(v_{l'}, v_b)=(-4.5
\mkms,-4 \mkms)$.  RSG-5 is visible at $(l,b)=(83^\circ,6^\circ)$,
$(v_{l'}, v_b)=(5.5 \mkms,-3.5 \mkms)$.  Most of the other
subclusters, including in Cep-Cyg ($l,b=90^\circ,7^\circ$) and
Cerberus ($l,b=48^\circ,18^\circ$) are too small or dispersed to have
previously been analyzed in great detail.

%
% KOI match numbers from calc_CepHer_group_numbers.py
%
Our fourth and final step was to cross-match the candidate Cep-Her
member list against all known Kepler Objects of Interest.  We used the
Cumulative KOI table from the NASA Exoplanet Archive from 27 March
2022, and also compared against the \texttt{q1\_q17\_dr25} table
\citep{thompson_planetary_2018}.  From the candidate members
with weights exceeding 0.02, this yielded
 yielded 11 were known false positives, 6 ``confirmed'' planets, and 8
 ``candidate'' planets.  
Inspection of the Kepler data validation summaries and Robovetter
classifications for these objects showed whether they were
potentially consistent with being {\it i)} planets, and {\it ii)}
$\lesssim 10^8$ years old, based on the presence of rotational
modulation at the expected period and amplitude \citep[{\it
e.g.},][Figure~9]{rebull_rotation_2020}.  Four objects remained after
this inspection: Kepler-1627, Kepler-1643, KOI-7368, and
KOI-7913.\footnote{These figures
in fact give an estimate of the size of Cep-Her: if our 4/14
possible matches represent the true positive rate, then
multiplying by the number of stars above our weight threshold
(12{,}436) yields $\approx$3{,}500 stars in the complex.}

Figure~\ref{fig:XYZvtang} shows the positions of the KOIs along
various projections.  Kepler-1643 is near the core RSG-5 population
both spatially and kinematically.  KOI-7368 and KOI-7913 are in a 
diffuse region $\approx$40\,pc above RSG-5 in $Z$ and $\approx$100\,pc
closer to the Sun in $Y$.  In tangential galactic velocity space,
there is some kinematic overlap between the region the
latter two KOIs are in and the main RSG-5 group.

%
% Numbers are from RSG-5_auto_XYZ_vl_vb_cut.csv,
% CH-2_auto_XYZ_vl_vb_cut.csv
%
We define two sets of stars in the local vicinity of our objects of
interest.  For candidate RSG-5 members, we require:
\begin{align}
  X/{\rm pc} &\in [45, 75] \nonumber \\
  Y/{\rm pc} &\in [320, 350] \nonumber \\
  Z/{\rm pc} &\in [40, 70] \nonumber \\
  v_b/{\rm km\,s^{-1}} &\in [-4, -3] \nonumber \\
  v_{l^*}/{\rm km\,s^{-1}} &\in [4, 6] \nonumber
\end{align}
For the diffuse stars near KOI-7368 and KOI-7913, we require
\begin{align}
  X/{\rm pc} &\in [20, 70] \nonumber \\
  Y/{\rm pc} &\in [230, 270] \nonumber \\
  Z/{\rm pc} &\in [75, 105] \nonumber \\
  v_b/{\rm km\,s^{-1}} &\in [-3.5, -1.5] \nonumber \\
  v_{l^*}/{\rm km\,s^{-1}} &\in [2, 6] \nonumber
\end{align}
and we call this latter set of stars ``CH-2''.  These cuts yield 141
candidate RSG-5 members, and 37 candidate CH-2 members.
Tables of these stars are enumerated in Appendix~\ref{app:members}.
An important consideration, especially for CH-2, is the
contamination rate by field stars.  We assess this in the following
section.

\subsection{The Cluster's Age}
\label{sec:clusterage}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2c.pdf}
			\includegraphics[width=0.49\textwidth]{f2d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf The stellar groups near KOI-7368, KOI-7913, and Kepler-1643
    are 40 to 50 million years old.} 
    {\it Top row}: 
    Color-absolute magnitude diagram of candidate Cep-Her members, plotted
    over candidate members of the $\delta$~Lyr~cluster
    ($\approx38$\,Myr; \citealt{bouma_kep1627_2022}) and the Gaia
    EDR3 Catalog of Nearby Stars (gray background).  The left and
    right columns shows stars in RSG-5 and CH-2, respectively.  The
    range of colors is truncated to emphasize the pre-main-sequence.
    Stars that fall far below the cluster sequences are field
    interlopers.
    {\it Bottom row}:
    TESS and ZTF-derived stellar rotation periods, with the Pleiades
    ($\approx 112$\,Myr) and Praesepe ($\approx 650$\,Myr) shown for reference
    \citep{rebull_rotation_2016a,douglas_poking_2017}.
    The detection efficiency for reliable rotation periods falls off
    beyond $\bpmrpo \gtrsim 2.6$.
	\label{fig:age}
	}
\end{figure*}

\subsubsection{Color-Absolute Magnitude Diagram}
\label{sec:camd}

Color-absolute magnitude diagrams (CAMDs) of the candidate RSG-5 and
CH-2 members are shown in the upper row of Figure~\ref{fig:age}.  The
stars from the $\delta$~Lyr cluster are from
\citet{bouma_kep1627_2022}, and the field stars are from the Gaia EDR3
Catalog of Nearby Stars \citep{gaia_gcns_2021}.  To make these
diagrams, we imposed the data filtering criteria from
\citet[][Appendix B]{GaiaCollaboration2018}, which 
include binaries while omitting instrumental artifacts from for
instance low photometric signal to noise, or a small number of
visibility periods.  We then corrected for extinction using the
\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}}
dust maps and the extinction coefficients $k_X\equiv A_X/A_0$ from
\citet{GaiaCollaboration2018}, assuming that $A_0 = 3.1 E(B-V)$.  This
yielded a mean and standard deviation for the reddening of
$E(B-V)=0.036\pm0.002$ for RSG-5, and $E(B-V)=0.017\pm0.001$
for CH-2.  By way of comparison, in \citet{bouma_kep1627_2022} the
same query for the $\delta$~Lyr cluster yielded
$E(B-V)=0.032\pm0.006$.  Finally, for the plots we set the $\bpmrpo$
color range to best visualize the region of maximal age information
content: the pre-main-sequence.

The CAMDs show that for RSG-5, almost all of our candidate members are
on a tight pre-main-sequence locus.  This implies a false positive
rate of a few percent, at most.  By comparison, our control sample
(the $\delta$~Lyr candidates) has a false positive rate of
$\approx$12\%, based on the number of stars that photometrically
overlap with the field population, rather than the bulk cluster
population.  For CH-2, our membership selection gives 27 objects in
the color range displayed, and 23 of them appear to be consistent with
being on the pre-main-sequence.  This implies a false positive rate in
CH-2 of $\approx$15\%.

Figure~\ref{fig:age} also shows that most RSG-5 and
CH-2 members overlap with the $\delta$~Lyr cluster, and that the groups are therefore
roughly the same age.  To quantify this, we use the empirical method
introduced by \citet[][their Section~6.3]{gagne_mutau_2020}.  This
idea is to fit the pre-main-sequence loci of a set of
reference clusters, and to then model the locus of the target cluster
as a linear combination of these reference cluster loci.  For our
reference clusters, we used UCL, IC\,2602, and the Pleiades, with the
memberships reported by \citet{Damiani2019} and
\cite{CantatGaudin2018a} respectively.  We adopted ages of 16\,Myr for
UCL \citep{pecaut_star_2016}, 38\,Myr for IC\,2602\footnote{ Ages for
IC\,2602 vary from 40 to 46\,Myr based on lithium-depletion-boundary
(LDB) measurements \citep{dobbie_ic_2010,randich_gaiaeso_2018}, and
from 30 to 46\,Myr based on isochronal analyses
\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.},
and 112\,Myr for the Pleiades \citep{dahm_2015}.  These assumptions
and the subsequent processing steps taken to exclude field stars and binaries
were identical to those
described in \citet{bouma_kep1627_2022}.  The mean and 
uncertainty of the resulting age posterior are $46^{+9}_{-7}$\,Myr for
RSG-5, and $36^{+10}_{-8}$\,Myr for CH-2.  For comparison, the this
procedure yields an age for the $\delta$~Lyr cluster of
$38^{+6}_{-5}$\,Myr.  
The slightly older isochronal age of RSG-5 is
consistent with its location
relative to the $\delta$~Lyr cluster in
the upper left panel of Figure~\ref{fig:age}.

\subsubsection{Stellar Rotation Periods}
\label{sec:rotation}

An independent way to assess the age of the candidate cluster members
is to measure their stellar rotation periods.  This approach can be
achieved using surveys such as TESS \citep{ricker_transiting_2015} and
the Zwicky Transient Facility (ZTF, \citealt{bellm_zwicky_2019});
it leverages a storied tradition of measuring rotation periods of stars in
benchmark open clusters \citep[see {\it
e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
our case are especially useful, since they provide 3 to 5 lunar months
of photometry for all of our candidate CH-2 and RSG-5 members.

We selected stars suitable for gyrochronology by requiring $\bpmrpo
\geq 0.5$ to focus on FGKM stars that experience magnetic braking.
For TESS, we also restricted our sample to $G<16$, to ensure the stars
are bright enough to extract usable light curves from the full-frame
images (FFIs).  The magnitude cut corresponds to $\bpmrpo < 2.6$
($\sim$M3V) at the relevant distances.  These cuts gave 19 stars in
CH-2 and 42 stars in RSG-5.  We extracted light curves from the TESS
images using the \texttt{unpopular} package
\citep{hattorio_2021_cpm}, and regressed them against systematics with
its causal pixel model.  We measured rotation periods using
Lomb-Scargle periodograms and visually vetted the results using an
interactive program that allows us to switch between TESS Cycles,
select particular sectors, flag stars with multiple periods, and
correct half-period harmonics. For ZTF, we used the same color cut to
focus on FGKM stars, but restricted the sample to $13 < G < 18$ to
avoid the saturation limit on the bright end and ensure sufficient
photometric precision at the faint end. We followed the procedure
outlined in \citet{curtis_rup147_2020}: we downloaded $8'\times8'$
image cutouts, ran aperture photometry for the target and neighboring
stars identified with Gaia, and used them to define a systematics
correction to refine the target light curves.

The lower panels of Figure~\ref{fig:age} show the results.  In RSG-5,
36/42 stars have rotation periods faster than the Pleiades (86\%).
This numerator omits the two stars with periods $>$$12$\,days visible in
the lower-left panel of Figure~\ref{fig:age}.  The age interpretation
for these latter stars, particularly the $\approx$M2.5 dwarf, is not
obvious.  \citet{rebull_usco_2018} for instance have found numerous
M-dwarfs with 10-12 day rotation periods at ages of USco ($\sim$$8$\,Myr),
and some may still exist at ages of LCC ($\sim$$16$\,Myr;
L.~Rebull in preparation).  Regardless, given that nearly no field
star outliers seem to be present on the RSG-5 CAMD, the fact that we
do not detect rotation periods for $\approx$14\% of stars should
perhaps be taken as an indication for the fraction of stars for
which rotation periods might not be detectable, due to {\it e.g.}, pole-on
stars having lower amplitude starspot modulation.

%
% CH-2
% 5 stars with Prot>10 days (from the Prot versus color plot):
%                      DESIGNATION  TESS_Data   bp_rp_x     period
%25  Gaia EDR3 2129930258400157440          4  1.909693  15.309000
%24  Gaia EDR3 2129417404945830784          4  2.040988  10.981582
%4   Gaia EDR3 2107169165113864064          4  2.059904  10.580494
%9   Gaia EDR3 2134851775526125696          4  2.326998  12.227900
%1   Gaia EDR3 2127726081184588160          4  2.347244  11.543281
%
% CH-2
% Photometric field stars (from Glue -- one of these is omitted due to phot S/N cut, 
% presumably 2126008609657702784 )
%
%In [9]: df[['DESIGNATION', 'bp_rp', 'M_G']]
%Out[9]:
%                     DESIGNATION     bp_rp        M_G
%0  Gaia EDR3 2126529468937923456  2.518077   9.844733 --> NaN period, G=16.9
%1  Gaia EDR3 2126008609657702784  2.746735  11.409737 --> NaN, G=18.4
%2  Gaia EDR3 2133330051428541952  2.499458   9.459546 --> 15.7 d (ZTF), G=16.5
%3  Gaia EDR3 2133510921093964928  2.720177  10.923209 --> NaN, G=17.9
%4  Gaia EDR3 2139369183468814464  2.929506  11.170221 --> NaN, G=18.2
%5  Gaia EDR3 2129930258400157440  1.909693   8.200652 --> P=15.3 (ZTF), G=15.2

For CH-2, 13/19 stars have rotation periods that are obviously faster
than their counterparts in the Pleiades (68\%).  4 stars, not included
in the preceding numerator, are M-dwarfs with rotation periods between
10 and 12.5 days.  As just discussed, the age interpretation for these
M-dwarfs is not obvious.  Regardless, the $\approx$15\% false positive
rate determined from the CAMD seems consistent with our fraction of
detected rotation periods, given that RSG-5 was also missing rotation
period detections for $\approx$15\% of its candidate members, which
all seemed photometrically consistent with being part of a single
pre-main-sequence locus.

It is challenging to convert these stellar rotation periods to a
precise age estimate, since on the pre-main-sequence the stars are
spinning up due to thermal contraction rather than down due to
magnetized braking.  Regardless, the rotation period distributions of
both CH-2 and RSG-5 seem consistent with other 30\,Myr to 50\,Myr
clusters ({\it e.g.}, IC\,2602 and IC\,2391;
\citealt{douglas_stephanie_t_2021_5131306}).  They also seem
consistent with the estimated false positive rates determined from the
color-absolute magnitude diagrams.

\section{The Stars}
\label{sec:stars}

\input{tab_sysparams.tex}

Some salient stellar properties of the Kepler objects in Cep-Her can
be gleaned from Figure~\ref{fig:age}.  They span spectral types of G8V
(Kepler-1627) to K6V (KOI-7913 A).  The secondary in the KOI-7913
system has spectral type $\approx$K8V.  And since a Solar-mass star
with solar metallicity arrives at the zero-age main sequence at
$t\approx40$ Myr \citep{choi_mesa_2016}, these stars are all in the
late stages of their pre-main-sequence contraction.  Their
heliocentric distances are $\approx$260\,pc (KOI-7913 and KOI-7368, in
CH-2) and $\approx$340\,pc (Kepler-1643, in RSG-5).

The adopted stellar parameters are listed in
Table~\ref{tab:sysparams}.  The stellar surface gravity, radius, mass,
and density are found by interpolating against the MIST isochrones
\citep{choi_mesa_2016}.  The statistical uncertainties from this
technique mostly originate from the parallax uncertaintes; the
systematic uncertainties are taken to be the absolute difference
between the PARSEC \citep{bressan_parsec_2012} and MIST isochrones.
Reported uncertainties are a quadrature sum of the statistical and
systematic components.

To verify these parameters and to also analyze spectroscopic youth
proxies such as the Li\,\textsc{I} 6708\,\AA\ doublet and H$\alpha$,
we acquired spectra.  We also acquired high resolution imaging for
each system, to constrain the existence of visual companions,
including possible bound binaries.  The system-by-system details
follow, and the relevant results are also summarized in
Table~\ref{tab:sysparams}.

\subsection{Kepler\,1643}

\paragraph{Spectra}
For Kepler-1643, we acquired two iodine-free spectra from Keck/HIRES
on the nights of 2020 Aug 16 and 2021 Oct 25.  The acquisition and
analysis followed the usual techniques of the California Planet Survey
\citep{howard_cps_2010}.  We derived the stellar parameters ($T_{\rm
eff}, \log g, R_\star$) using \texttt{SpecMatch-Emp}
\citep{yee_SM_2017}, which yielded values in $<1$-$\sigma$ agreement
with those from the cluster-isochrone method.  This approach also
yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.  Using the broadened synthetic
templates from \texttt{SpecMatch-Synth} \citep{petigura_cksi_2017}, we
found $v\sin i = 9.3 \pm 1.0\,\mkms$.  The systemic radial velocity at
the two sequence epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm
1.2\,\mkms$ respectively.  To infer the equivalent width of the
Li\,\textsc{I} 6708\,\AA\ doublet, we followed the procedure described
by \citet{bouma_2021_ngc2516}.  This yielded a strong detection: ${\rm
EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with values consistent at
$<1$-$\sigma$ between the two epochs.   The quoted value does not
correct for the Fe\,\textsc{I} blend at 6707.44\AA, because HIRES
resolves the two features.

\paragraph{High-Resolution Imaging}
We acquired adaptive optics imaging of Kepler-1643 on the night of
2019 June 28 using the NIRC2 imager on Keck-II.  Using the narrow
camera (FOV = 10.2\arcsec), we obtained 4 images in the $K'$ filter
($\lambda = 2.12\,\mu$m) with a total exposure time of 320\,s. We
analyzed these data following \citet{kraus_impact_2016}, and
determined the detection limits to visual companions by analyzing the
residuals after subtracting an empirical PSF template.  This procedure
yielded contrast limits of $\Delta K' = 4.1$ mag at $\rho = 150$ mas,
$\Delta K' = 5.8$ mag at $\rho = 300$ mas, and $\Delta K' = 8.3$ mag
at $\rho > 1000$ mas.

\subsection{KOI-7368}
\paragraph{Spectra}
For KOI-7368, we acquired a spectrum on 2015 June 1 using the echelle
spectrograph (TRES; \citealt{furesz_tres_2008}) mounted at the
Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.  The
stellar parameter classification pipeline for the TRES spectra has
been described by \citet{2021tsc2.confE.124B}.  It is based on the
synthetic template library originally constructed by
\citet{buchhave_hatp16b_class_2010}.  The resulting stellar parameters
($T_{\rm eff}, \log g, R_\star$) agreed with those from the
cluster-isochrone method within $1$-$\sigma$, though the effective
temperature was more precise by a factor of three.  Auxiliary
spectroscopic parameters included the metallicity $[{\rm Fe/H}]= -0.02
\pm 0.08$, the equatorial velocity $v\sin i = 20.21 \pm 0.50\,\mkms$,
and the systemic velocity ${\rm RV}_{\rm sys} = -10.9 \pm 0.2\,\mkms$.
The Li 6708\AA\ EW measurement procedure yielded ${\rm EW}_{\rm Li} =
236^{+16}_{-14}$\,m\AA.

\paragraph{High-Resolution Imaging}
We acquired adaptive optics imaging of KOI-7368 on the night of 2019
June 12, again using NIRC2.  The observational configuration and
reduction was identical as for Kepler-1643.  No companions were
detected, and the analysis of the image residuals yielded contrast
limits of $\Delta K' = 5.2$ mag at $\rho = 150$ mas, $\Delta K' = 6.7$
mag at $\rho = 300$ mas, and $\Delta K' = 8.7$ mag at $\rho > 1000$
mas.

\subsection{KOI-7913}

\paragraph{Binarity}
KOI-7913 is almost certainly a binary.  The north-west primary is
$\approx$0.5 magnitudes brighter than the south-east secondary in
optical passbands.  The two stars are separated in Gaia EDR3 by
$3\farcs5$ on-sky, and have parallaxes consistent within $1$-$\sigma$
(with an average $\varpi=3.66 \pm 0.01$\,mas).  The apparent on-sky
separation is $959 \pm 2$ au.  The Gaia EDR3 proper motions are also
nearly identical.  Since two stars were resolved in the Kepler Input
Catalog and are roughly one Kepler pixel apart, an accurate crowding
metric has already been applied in the NASA Ames data products to
correct the mean flux level \citep{2017ksci.rept....6M}.  This is
important for deriving accurate transit depths.

\paragraph{Spectra}
We acquired Keck/HIRES spectra for KOI-7913 A on the night of 2021 Nov
13, and KOI-7913 B on the night of 2021 Oct 26.  The
\texttt{SpecMatch-Emp} \citep{yee_SM_2017} machinery yielded $T_{\rm
eff,A} = 4324 \pm 70\,{\rm K}$, $T_{\rm eff,B} = 4038 \pm 70\,{\rm
K}$.  The remaining parameters were in agreement with those from the
cluster isochrone.  For the primary, we also found $[{\rm Fe/H}]=
-0.06 \pm 0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm
sys} = -17.8 \pm 1.1\,\mkms$.  For the secondary, these same
parameters were $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm
1.0\,\mkms$, and ${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.  Neither
of the KOI-7913 components shows strong lithium absorption, but this
is expected given their $\approx$K6V and $\approx$K8V spectral types.
They do however both show H$\alpha$ in emission.  We discuss the
significance in greater depth in Appendix~\ref{app:spectra}, but
broadly speaking both observations are consistent with a $\approx$40
Myr age for KOI-7913.

\paragraph{High-Resolution Imaging}
We acquired adaptive optics imaging of KOI-7913 on the night of 2020
Aug 27 using the NIRC2 imager.  The observational configuration and
reduction were identical as before.  The images showed KOI-7913 A,
KOI-7913 B, and an additional faint star $\approx$$0\farcs99$ due East
of KOI-7913 B.  Applying the PSF-fitting routines from
\citet{kraus_impact_2016}, the tertiary star has a separation $\rho =
4397 \pm 3$\,mas from the primary, at a position angle $231.17^\circ
\pm 0.02^\circ$, with $\Delta K' = 6.97 \pm 0.04$.  While it is too
faint to affect the interpretation of the KOI itself, it would be
amusing if the faint neighbor were also comoving and young -- at the
age of the system, it would have a mass roughly between 10 and
15\,M$_{\rm Jup}$.  Additional imaging epochs will tell.

\section{The Planets}
\label{sec:planets}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.85\textwidth]{f3e.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.33\textwidth]{f3a.pdf}
			\includegraphics[width=0.33\textwidth]{f3b.pdf}
		}

		\vspace{-1.27cm}	
		\subfloat{
			\includegraphics[width=0.33\textwidth]{f3c.pdf}
			\includegraphics[width=0.33\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Raw and processed light curves for the Kepler Objects of
    Interest in Cep-Her.}   
    {\it Top}:
    50 day light curve segment from the 3.9 years of Kepler data.  The
    ordinate shows the \texttt{PDCSAP} median-subtracted flux in units
    of parts-per-thousand ($\times 10^{-3}$).  The dominant signal is
    induced by starspots.  
    {\it Bottom}:
    Phase-folded transits of Kepler-1643, KOI-7913, KOI-7368, and
    Kepler-1627 with stellar variability removed.  The maximum {\it a
    posteriori} transit model is shown with the gray line, and the
    residual after subtracting the transit model is vertially
    displaced.  Windows over 10 hours are shown.  Gray points are
    individual flux measurements; black points are binned to 20 minute
    intervals, and have a representative 1-$\sigma$ error bar in the
    center-right of each panel. 
		\label{fig:planets}
	}
\end{figure*}

\subsection{Kepler Data}

The Kepler space telescope observed Kepler-1643, KOI-7913, and
KOI-7368 at a 30-minute cadence between May 2009 and April 2013.  For
all three systems quarters 1 through 17 were observed with minimal
data gaps.  The top panel of Figure~\ref{fig:planets} shows a 50-day
slice of the \texttt{PDCSAP} light curves for the three new Cep-Her
candidates, along with Kepler-1627.  In \texttt{PDCSAP},
nonastrophysical variability is removed through a cotrending approach
that uses a set of basis vectors derived by applying singular value
decomposition to a set of systematics-dominated light curves
\citep{smith_kepler_PDC_2017}.  In our analysis, we used the
\texttt{PDCSAP} light curves with the default optimal aperture
\citep{smith_finding_2016}.  Cadences with non-zero quality flags were
omitted.  In all cases, the stars are dominated by spot-induced
modulation with peak-to-peak variability between 2\% and 10\%.  These
signals are much larger than the transits, which have depth
$\approx$0.1\%.  To quantify the stellar rotation periods, we
calculated the Lomb-Scargle periodogram for each Kepler quarter
independently.  The resulting means and standard deviations are in
Table~\ref{tab:sysparams}.

\subsection{Transit and Stellar Variability Model}
\label{sec:fitting}

Our goals in fitting the Kepler light curves are twofold.  First, we
want to derive accurate planetary sizes and orbital properties.
Second, we want to remove the stellar variability signal to enable a
statistical assessment of the probability that the transit signals are
planetary.

Our adopted approach was as follows.  Given an initial guess for the
transit ephemeris from the \texttt{q1\_q17\_dr25} analysis
\citep{thompson_planetary_2018}, the light curve was first trimmed to
a local window around each transit, spanning $\pm 3\times t_{14}$.
The local out-of-transit points were then fitted with a fourth-order
polynomial, which was then divided out from the light curve.  The
resulting flattened transits were then fitted with a planetary transit
under the assumption of quadratic limb darkening.  Our model therefore
included 8 free parameters for the transit ($\{P, t_0, \log R_{\rm
p}/R_\star, b, u_1 ,u_2, R_\star, \log g\}$), 2 free parameters for
the light curve normalization and a white noise jitter ($\{\langle f
\rangle, \sigma_f \}$), and an additional 5 implicit parameters per
transit through the polynomial.

We fitted the models using \texttt{exoplanet}
\citep{exoplanet:exoplanet}.  We assumed a Gaussian likelihood, and
sampled using \texttt{PyMC3}'s No-U-Turn Sampler
\citep{hoffman_no-u-turn_2014}, after having initialized using the
maximum {\it a posteriori} (MAP) model.  We used the
\citet{gelman_inference_1992} statistic ($\hat{R}$) as our convergence
diagnostic.  The resulting fits are shown in the lower panels of
Figure~\ref{fig:planets}, and the important derived parameters are in
Table~\ref{tab:sysparams}.  The set of full parameters with their
priors are given in Appendix~\ref{app:transit}.

One weakness of our fitting approach is that we have fixed 5 implicit
parameters per individual transit to their MAP values to remove the
stellar variability signal.  An alternative could be to fit the
planetary transits simultaneously with the starspot-induced
variability using, say, a quasiperiodic Gaussian process (GP).  We
explored this approach, but found that it required careful fine-tuning
of the hyperparameter priors, otherwise the GP would tend to
incorporate variability that was not part of the transit, even if it
was short-term variability attributable to flares, or simply
instrumental noise.  This failure mode is somewhat pernicious, in that
it yields an ill-founded sense of confidence in an ``overly clean''
transit model fit.  While our model is simpler, it has the benefit
that the white noise jitter never trades off with any parameter
equivalent to a damping timescale for the coherence of the GP.  It is
also computationally efficient, and it captures the planetary
parameters about which we care the most.

\subsection{Planet Validation}

In the future, it may be possible to obtain independent evidence for
the Cep-Her planets to directly confirm their planetary nature.  The
most likely path would be through transit observations using a
high-resolution spectrograph.  In
the interim, it is of interest whether the transit signals could be
astrophysical false positives, or whether they can be argued to be
planetary on a statistical basis.  We adopt the Bayesian framework
implemented in \texttt{VESPA} for this purpose
\citep{morton_efficient_2012,vespa_2015}.  Briefly summarized, the
priors in \texttt{VESPA} take as given the binary star occurrence rate
from \citet{raghavan_survey_2010}, direction-specific star counts from
\citet{girardi_star_2005}, and planet occurrence rates as described by
\citet[][Section~3.4]{morton_efficient_2012}.  The likelihoods are
then evaluated by forward-modeling a synthetic population of eclipsing
bodies for each possible astrophysical model class, in which each
population member has a particular trapezoidal eclipse depth, total
duration, and ingress duration.  These summary statistics are then
compared against the actual photometric data to evaluate the
probabilities of false positive scenarios such as foreground eclipsing
binaries, hierarchical eclipsing binaries, and background eclipsing
binaries.

\paragraph{Kepler-1643}
Kepler-1643~b (KOI-6186.01) was already validated as a transiting
planet by \citet{morton_false_2016}, who found a probability for any
of the aforementioned false positive scenarios of 9$\times$10$^{-6}$.
Repeating the calculation with our own stellar-variability correction
and the new NIRC2 imaging constraints, we find ${\rm FPP} =
6\times10^{-9}$, where the improvement is due to the new NIRC2
contrast curves.  Figure~\ref{fig:planets} shows the justification:
the transit is flat and has a high S/N ($\approx 47$).  The shape is
therefore difficult to reproduce with eclipsing binary models.

Intriguingly, Kepler-1643 formally failed one of the centroid shift
tests (see the \texttt{q1\_q17\_dr25\_koi} data release): the angular
distance between the target star's catalog position and the position
of the transiting source was measured as $1\farcs0$ at 4.4-$\sigma$.
The reports show however that two outlying quarters (2 and 6) drive
the offset -- the centroid locations from the other Kepler quarters
are consistent at $\lesssim 0\farcs4$.  Nonetheless, this is an
instructive exercise in how stellar variability complicates
centroid-based vetting tests.  The centroid shifts measured by these
tests are determined from the in- and out-of-transit flux-weighted
centroids.  For stars with significant spot-induced variability there
is no static baseline in either the in- or out-of-transit phases.

\paragraph{KOI-7368}
KOI-7368.01 is listed on the NASA Exoplanet Archive as a ``candidate''
planet.  \citet{morton_false_2016} did not compute a false positive
probability for the system because their default trapezoidal fitting
routine failed, presumably due to the spot-induced variability.  Our
fitting approach rectifies this point, and our new NIRC2 images
revealed no new stellar companions.  Performing the relevant
calculation, we find ${\rm FPP} \approx 4\times10^{-3}$.  Though not
as convincing as Kepler-1643, this formally clears the usual threshold
for calling the planet statistically validated
\citep{morton_efficient_2012}.  The S/N of the transit is
$\approx$$32$, which indicates that it is unlikely to be caused by
systematic noise in the light curve (see Figure~\ref{fig:planets}).
The positional probability score also meets the requirements for
transiting sources thought to share positions with their target stars
\citep{2017ksci.rept...16B}.

It bears mentioning that KOI-7368 shows a weak centroid shift in the
\texttt{q1\_q17\_dr25\_koi} validation reports, similar to
Kepler-1643.  For KOI-7368, the reported offset is smaller, and less
statistically significant ($0\farcs2$; 3.0-$\sigma$).  Again, the data
validation reports show that the shift is caused by a few outlying
quarters (4, 5, 8, and 12).  The fact that the remaining data show a
consistent scatter between centroid locations suggests that these
outliers are also caused by the stellar variability.  Independently,
our NIRC2 imaging shows that there are no neighboring sources that
could cause an offset of the observed amplitude.

\paragraph{KOI-7913}
KOI-7913.01 is also currently listed on the NASA Exoplanet Archive as a
``candidate'' planet.  The canonical \citet{morton_false_2016}
analysis was of Q1-Q17 KOIs from DR24, and therefore spanned KOI-1.01
to KOI-7620.01 and omitted KOI-7913.01.  However the results of the
subsequent DR25 analysis by Morton et al.\ are listed at the NASA
Exoplanet Archive.  The relevant table gives a probability for the
system being an astrophysical false positive of $1.4\times10^{-4}$,
with the most likely false positive scenario being a blended eclipsing
binary.  Repeating the calculation with our new detrending and
contrast curves, we find a similar result: ${\rm FPP} =
1.3\times10^{-4}$.  Though the transit has the lowest S/N of any of
the objects discusssed (${\rm S/N}=14$), its lower FPP relative to
KOI-7368 can be understood through its relatively flat-bottomed shape,
combined with its long transit duration relative to plausible
eclipsing binary models (Figure~\ref{fig:planets}).  The host star
probability usability score \citep{2017ksci.rept...16B} also meets the
usual threshold, and so the planet seems to be validated.  Its
disposition has however previously fluctuated from ``false positive''
to ``candidate''.  The most likely explanation is the presence of
KOI-7913~B, which is located approximately one Kepler pixel away from
Kepler-7913 A (see Appendix~\ref{app:koi7913}).  Our pixel-level
analysis of the target pixel files did not show any evidence to
support the idea that KOI-7913~B could host the transiting planet.
Our NIRC2 imaging similarly ruled out the possibility of any other
plausible host star for the planet other than KOI-7913~A.

\section{Discussion \& Conclusion}
\label{sec:disc_conc}

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		%    namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'KOI-7913 B', 'Kepler-1643']
		%    markers = ['P', 'v', 'X', 'X', 's']
		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
    3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
    its uncertainty. Kepler-1627 (+), KOI-7368 (down-triangle),
    KOI-7913 (X), Kepler-1643 (diamond).  The large sizes of the
    youngest transiting planets could be tied to a selection effect.
    Kepler-1643, KOI-7368, and KOI-7913 have normal mini-Neptune sizes
    of 2 to 3\,$R_\oplus$, which is a novelty given their ages.
    Parameters are from the NASA Exoplanet Archive (2022 Apr 5).
		\label{fig:rp_period_age}
	}
\end{figure*}

\subsection{Normal-Sized Mini-Neptunes Exist at 40$\,$Myr}
\label{subsec:sizes}

The most significant novelty about
Kepler-1643~b, KOI-7368.01, and KOI-7913~Ab is that their sizes are
normal, relative to the known population of gigayear-old
mini-Neptunes.  At field star ages, mini-Neptune sizes span
1.8\,$R_\oplus$ to 3.6\,$R_\oplus$, with the most
common size being $\approx 2.4\,R_\oplus$
\citep[][Figure~7]{Fulton_et_al_2017}.  The known planets younger than
$10^8$ years are almost all larger, with sizes between 4 and
10$\,R_\oplus$
\citep{Mann_K2_33b_2016,David_et_al_2017,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020,mann_toi1227_2022}.
Figure~\ref{fig:rp_period_age} explores this in more detail by showing
the sizes, orbital periods, and ages of the known transiting planets,
emphasizing planets with precise ages.  The smallest previously known planets
comparable to the new Cep-Her mini-Neptunes
are AU~Mic~c ($3.0\pm0.2\,R_\oplus$, see
\citealt{martioli_aumicbc_2021} and \citealt{gilbert_flares_2022}),
Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$; \citealt{bouma_kep1627_2022}),
and AU~Mic~d ($4.2\pm0.2\,R_\oplus$; \citealt{plavchan_planet_2020}).

The theoretical expectation is that mini-Neptunes with sizes of 2 to
3\,$R_\oplus$ should be common at ages of $10^7$ to $10^8$ years.
This expectation is tied to inferences about the
initial distributions of planetary core mass, core composition, and
atmospheric mass fraction \citep{owen_evaporation_2017}.  The
Kelvin-Helmholtz cooling timescale, which is tied to the entropy of
the planetary interior once the disk disperses, could also play a
significant role \citep{owen_constraining_2020}.  As an example,
\citet{rogers_unveiling_2021} predicted that given a core-mass
distribution peaked at $\approx$$4$\,M$_\oplus$, an ice-poor rock/iron
core composition, and a typical H/He mass fraction of $\approx$4\%,
there should be a maximum in planet occurrence at 2 to 3\,$R_\oplus$,
at times between 10 and 100\,Myr
(their Figure~16).  The models advanced by
\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} agree;
the theoretical differences lie in their explanations
for the origin of the radius valley.  If the radius gap is primordial
and created by gas accretion physics, then super-Earths
($<$$1.8$\,$R_\oplus$) would be common as early as $10^7$ years
\citep{lee_primordial_2021}.   The photoevaporation-driven models more
often quote $10^8$ year timescales for production of
atmosphere-stripped super-Earths, while the models with the mass-loss
driven by a slow escape of heat from the core predict longer
$\sim$$10^9$ year timescales (but see \citealt{king_euv_2020}).

Systems such as K2-25, V1298~Tau, HIP-67522, TOI-837,
and TOI-1227 have sizes that are much larger than the picture we have
just described.  However, their large sizes can be accommodated
by existing models by invoking any of {\it i)} larger core masses,
{\it ii)} more volatile-rich compositions, {\it iii)}  larger initial
atmospheric mass fractions, or {\it iv)} longer thermal cooling times.
For instance, perhaps all five of the the aforementioned systems are
large simply because the planets are more massive than most
mini-Neptunes.  Secure mass measurements would help constrain this
parameter space, but the $\sim$1\,\kms\ spot-induced radial velocity
semi-amplitudes makes measuring the Doppler orbits incredibly
challenging.  \citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
Regardless, the new Kepler-1643, KOI-7368, and KOI-7913 systems do
demonstrate that at least some planets at 40\,Myr have sizes that are
consistent with theoretical expectations for mini-Neptunes.  While
selection effects imposed by spot-induced photometric variability are
a likely explanation for why planets this small have not previously
been identified \citep[{\it e.g.},][]{zhou_2021_tois}, future work
should quantify this bias more carefully, in order to enable empirical
studies of how the planetary size distribution changes at early times.

\subsection{Is CH-2 a real cluster?}
\label{subsec:ch2}

RSG-5, and Kepler-1643's membership inside it, meet typical
expectations for a star claimed to be in an open cluster.  RSG-5 is an
obvious overdensity relative to the local field, and our
membership selection easily produces a clean pre-main-sequence locus
in color versus absolute magnitude (Figure~\ref{fig:age}).  CH-2, and
KOI-7913 and KOI-7368's membership inside it, do not meet these
expectations in as obvious a manner.  This is because the CH-2
association is diffuse.

To quantify the density difference between CH-2 and RSG-5, we can compare the spatial and
velocity volumes searched for each group.
For RSG-5, we drew 141 candidate members from a $30\,{\rm
pc}\times30\,{\rm pc}\times30\,{\rm pc}$ spatial cube, given a $1 {\rm
\mkms} \times 2 {\rm \mkms }$ rectangle in apparent galactic velocity.
For CH-2, our 37 candidate members came from a rectangular prism of
dimension $50\,{\rm pc}\times40\,{\rm pc}\times30\,{\rm pc}$, and a
rectangular box of $2 {\rm \mkms} \times 4 {\rm \mkms}$.  If we define
the searched volume in units of ${\rm pc}^3\,{\rm km^2}\,{\rm s^{-2}}$, then
the volume ratio of CH-2 to RSG-5 is 9 to 1.  The number of candidate
members per unit searched volume in RSG-5 relative to CH-2 is
therefore 34 to 1.

Given its low density, in CH-2 really a star cluster?  For this
discussion, let us adopt the definition that a star cluster is a group
of at least 12 stars that was physically associated at its time of
formation.  The somewhat arbitrary ``12'' is set to distinguish star
clusters from high-order multiples \citep[see][]{krumholz_star_2019}.
Dissolved clusters and their tidal tails are therefore included in our
concept of clusters.  We do not require however that a particular number of stars per
unit spatial volume is required for a group to be a ``cluster''.
The latter point acknowledges that an important
factor in cluster identification is also the density in velocity
space.  Perhaps once stellar rotation periods and chemical abundances
reach the same level of ubiquity as stellar proper motions, they might
enable further refinement of our ability to discover stars that formed
together.

From a data-driven perspective, demonstrating that a group of stars
was physically associated at its time of formation is challenging.
While some young groups show kinematic evidence for expansion
\citep{kuhn_kinematics_2019}, many, including Sco-Cen, do not
\citep{wright_kinematics_2018}.  This complicates the feasibility of
deriving kinematic ages through traceback, as
well as through the expansion itself
\citep[see][]{crundall_chronostar_2019}.  A more minimal approach is
that suggested by \citet{tofflemire_2021}: search for coeval,
phase-space neighbors, measure their ages, and determine if they share
a common age.  This approach can demonstrate whether a star is {\it
currently associated} with a set of coeval stars, which is much easier
than determining what the association looked like in the past.  Our
analysis of CH-2 meets the latter standard for the existence of a
$\approx$40\,Myr assocation today.  It would be a worthy exercise for
future work to perform a similar search for coeval phase-space
neighbors on the entire dataset of known exoplanet hosts.  For the
time being, we can offer the anecdotal point that in our experience,
most stars do not have dozens of $\sim$40\,Myr neighbors within a
local volume of a few \kms\ and tens of parsecs.

\subsection{Future work}

\subsubsection{Cep-Her}
\paragraph{Kinematics}
* Need an RV survey.

\paragraph{Age(s)}
Benchmarking its age(s):
* Lithium analysis of FGK depletion (bright enough to be feasible).

* Asteroseismology of the delta Sct stars (TESS data are there, group is
rich enough).

\paragraph{More objects}

* TESS data: giant planets in Cep-Her.  We need more 1e7-1e8 year old
stars for the HJ occurrence versus time question.  This will help.

* Kepler data: exocomets?  Search for smaller planets?

\subsubsection{Occ rates}
* Kepler data... presumably $\sim$200 stars in the cluster were
observed, in order to have detected 4 planets. 
Occurrence rates?

\subsection{Summary}

We have argued that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab {\it i)}
are 40 to 50 million years old, and {\it ii)} that each signal
is most likely planetary.  The evidence for the planetary
interpretation comes from an application of \texttt{VESPA} to the
Kepler data, alongside new adaptive optics imaging from NIRC2.  The
validity of the \texttt{VESPA} framework rests on the premise that
non-astrophysical false positives, including false alarms and known
neighboring eclipsing binaries, can be rejected.  Kepler-1643 and
KOI-7368 both pass all but one of the canonical vetting checks: the
check for centroid offsets between the in and out-of-transit phases.
For both cases, we have argued the weak observed shifts seem to be
caused by starspot-induced variability in specific quarters
shifting the stellar center-of-light.
In addition, our imaging rules out companions stars with the necessary
brightness and position to explain the reported shifts.

Each system has multiple indicators of youth that support the
reported ages.  For Kepler-1643, perhaps the strongest youth indicator
is its physical and kinematic association with RSG-5.  Based on the
color-absolute magnitude diagram, we are able to select members
of this cluster with a false positive rate of $\ll 5\%$
(Figure~\ref{fig:age}); Kepler-1643 is one such member.  While the stellar
rotation period period agrees with this assessment, its lithium
content is somewhat low, which might motivate future exploration
of lithium depletion across FGKM stars in RSG-5 (see Appendix~\ref{app:spectra})

The spatio-kinematic argument for the youth of KOI-7368 and
KOI-7913 is weaker because they are in an association of stars, CH-2,
that is more diffuse.  For KOI-7913, stronger indicators of its age
come from its binary.  Both stellar components in KOI-7913
have isochronal ages consistent with $40$\,Myr.
Both components also show H$\alpha$ in emission, which for
the $\approx$K6V primary is a good indicator that the star is
$\lesssim$100$\,$Myr old.  For KOI-7368, which is slightly more
massive, the Li 6708\AA\ EW measurement and stellar rotation period
are more constraining, and provide independent verification of the
star's youth.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\clearpage
\acknowledgements
\raggedbottom
%
L.G.B{.} acknowledges support from the TESS GI Program (NASA grants
80NSSC19K0386 and 80NSSC19K1728) and the Heising-Simons Foundation (51 Pegasi~b
Fellowship).
% programs G011103 and G022117, through 
%
%
%FIXME
Keck/NIRC2 imaging was acquired by program 2015A/N301N2L
(PI: A.~Kraus). % and 2019A/N069 (PI: E.~Petigura).
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%
%
%This study was based in part on observations at Cerro Tololo
%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
%Association of Universities for Research in Astronomy (AURA) under a
%cooperative agreement with the National Science Foundation.
%
%
%Finally, this research has made use of the Keck Observatory Archive (KOA),
%which is operated by the W. M. Keck Observatory and the NASA Exoplanet
%Science Institute (NExScI), under contract with the National
%Aeronautics and Space Administration.
Finally, we also thank the Keck Observatory staff for their support of
HIRES and remote observing.  We recognize the importance that the
summit of Maunakea has within the indigenous Hawaiian community, and
are deeply grateful to have the opportunity to conduct observations
from this mountain.
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

% %
% This research made use of the NASA Exoplanet Archive, which is
% operated by the California Institute of Technology, under contract
% with the National Aeronautics and Space Administration under the
% Exoplanet Exploration Program.
% %

% Resources supporting this work were provided by the NASA High-End
% Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
% Division at Ames Research Center for the production of the SPOC data
% products.
%

% A.J.\ and R.B.\ acknowledge support from project IC120009 ``Millennium
% Institute of Astrophysics (MAS)'' of the Millenium Science Initiative,
% Chilean Ministry of Economy. A.J.\ acknowledges additional support
% from FONDECYT project 1171208.  J.I.V\ acknowledges support from
% CONICYT-PFCHA/Doctorado Nacional-21191829.  R.B.\ acknowledges support
% from FONDECYT Post-doctoral Fellowship Project 3180246.
% %
% C.T.\ and C.B\ acknowledge support from Australian Research Council
% grants LE150100087, LE160100014, LE180100165, DP170103491 and
% DP190103688.
% %
% C.Z.\ is supported by a Dunlap Fellowship at the Dunlap Institute for
% Astronomy \& Astrophysics, funded through an endowment established by
% the Dunlap family and the University of Toronto.
% %
% D.D.\ acknowledges support through the TESS Guest Investigator Program
% Grant 80NSSC19K1727.
%
%
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

\software{
  %\texttt{arviz} \citep{arviz_2019},
  %\texttt{altaipony} \citep{ilin_flares_2021},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  %\texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  %\texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
   	exoplanet:theano},
	%\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  %\texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  %\texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  %\texttt{numpy} \citep{walt_numpy_2011}, 
  %\texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  %\texttt{TESS-point}  \citep{burke_2020},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  %\texttt{wotan} \citep{hippke_wotan_2019}.
}
\

\facilities{
 	{\it Astrometry}:
 	Gaia \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2021_edr3}.
 	{\it Imaging}:
    Second Generation Digitized Sky Survey. %,
    %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
 	Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	%Gemini:North~(`Alopeke; \citealt{scott_nessi_2018,scott_twin_2021}.
 	{\it Spectroscopy}:
	%CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
  %PFS ({\bf CITE}),
	Tillinghast:1.5m~(TRES; \citealt{furesz_tres_2008}).
  %MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
	%AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
 	Keck:I~(HIRES; \citealt{vogt_hires_1994}).
 	%VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
%	  ASTEP:0.40$\,$m (ASTEP400),
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
%	  El Sauce:0.356$\,$m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
	  Kepler \citep{borucki_kepler_2010},
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
%   MuSCAT3 \citep{Narita_2020},
% 	NTT,
% 	SOAR (SOI),
 	  TESS \citep{ricker_transiting_2015}.
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
}

% \input{starparams.tex}
% \input{model_posterior_table.tex}
% \input{t1_appendix.tex}
% \input{t2_appendix.tex}

\clearpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\appendix
\section{Table of Candidate RSG-5 and CH-2 Members}
\label{app:members}

{\bf LGB TODO: add Gaia source lists and abbreviated info.  Include rotation periods.}

We measured rotation periods for 53 stars with ZTF and 51 stars with
TESS, for a total of 71 periods reported for unique stars and 32
overlapping between the surveys. We adopted the ZTF period over the
TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the
Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF
period of 6.61 days; however, we flagged it as a candidate
double-dipper, which appears to have inaccurately doubled the TESS
period to 13.34 days; (2) Gaia EDR3 2081737529891330560: we found 3.06
days with TESS and 6.64 days with ZTF; we suspect that TESS captured
the 1/2-period harmonic and adopt the approximately double value from
ZTF; (3) 2134851775526125696: for this star, we measured 1.91 days
with TESS from Cycle 2, but noted that the signal appeared to be
missing in Cycle 4; ZTF found a strong signal at 12.23 days and we
adopt this as the star's period. In the remaining overlap cases, we
adopted the average between TESS and ZTF as the final period. For
these overlap stars, the median absolute deviation is 0.01 days,
showing remarkable consistency between the surveys. For three stars,
we failed to detect a period in TESS but recovered one from ZTF; in
all cases the periods appear to be 13--16 days, pushing the limit of
what is detectable with TESS with our procedure: (1) Gaia EDR3
2129930258400157440, TESS showed a flat light curve while ZTF yielded
a 15.3-day period; (2) Gaia EDR3 2082376861542398336, LS found a
7.6-day period which we rejected during visual validation; we found
15.4 days with ZTF, and we suspect that the weak/rejected singal form
TESS might have been a 1/2 period harmonic; (3) Gaia EDR3
2082397099429013120, similar to the previous case, we rejected a
6.7-day signal from TESS and recovered a 12.8-day period with ZTF.

\section{Spectroscopic Youth Indicators}
\label{app:spectra}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.99\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.3cm}
	\caption{
    {\bf Spectroscopic youth diagnostics for Kepler-1627, KOI-7368,
    Kepler-1643, and KOI-7913 AB. }
    The spectra are shown in the observed frame, and
    the stars are sorted left-to-right in order of decreasing effective temperature.
    \label{fig:koiyouthindicators}
	}
\end{figure*}

Figure~\ref{fig:koiyouthindicators} shows key portions of the HIRES
and TRES spectra for the Kepler objects in Cep-Her.
Lithium absorption is obvious at 6708\AA\ in all stars except KOI-7913 B, where it is marginal.
H$\alpha$ is in emission for both components of KOI-7913,
and in absorption for the hotter stars.
In the following, 
we compare these observations against stars in benchmark
open clusters.

\subsection{Lithium}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f6a.pdf}
			\includegraphics[width=0.49\textwidth]{f6b.pdf}
		}

		\subfloat{
			\includegraphics[width=0.49\textwidth]{f6c.pdf}
		}
	\end{center}
	\vspace{-0.3cm}
	\caption{
    {\bf Lithium 6708\AA\ and H$\alpha$ equivalent widths for the
    objects of interest compared to young open clusters and
    field stars. }
	Positive equivalent width means absorption; negative
	equivalent width means emission.
    The statistical uncertainties on the newly measured equivalent
    widths are shown, or else are smaller than the markers.
    {\it Top}:
    The field stars are stars with Kepler planets from \citet{berger_identifying_2018}.
    The ``40-50 Myr'' reference stars (left) are from
		IC\,2602 \citep{randich_gaiaeso_2018} and Tuc-Hor
		\citep{kraus_stellar_2014}.
    The ``112 Myr'' stars (right) are from the Pleiades
    \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018}.
    {\it Bottom}:
    The H$\alpha$ comparison is against Tuc-Hor
    \citep[$\approx$$40$\,Myr;][]{kraus_stellar_2014}
    \label{fig:lithium}
	}
\end{figure*}

Figure~\ref{fig:lithium} compares the measured lithium equivalent
widths of the Kepler objects against a few reference populations.
We selected studies from the literature only where
upper limits were explicitly reported.  KOI-7368 and
KOI-7913~A have secure lithium detections, and the KOI-7913~B
detection is marginal (${\rm EW}_{\rm Li} = 42^{+12}_{-19}$).  For all
three stars, as well as for Kepler-1627~A, the observed lithium
equivalent width is consistent with the stellar effective temperatures
and $\approx40$\,Myr ages.

Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr sequence
in the left-panel of Figure~\ref{fig:lithium}, while still being above
the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$).
The right panel shows the comparison against the Pleiades, where
Kepler-1643 is more consistent with the observed dispersion in lithium.

One explanation for this could be that Kepler-1643 is a field
interloper; another could be that RSG-5 is older than 50\,Myr.  We do
not favor the former, given that {\it i)} the star was selected based
on its spatio-kinematic proximity to other RSG-5 members, nearly all
of which appear consistent with being $\approx$50\,Myr old
(Figure~\ref{fig:age} upper left), and {\it ii)} the same RSG-5
members display a gyrochronal sequence consistent with being younger
than the Pleaides (Figure~\ref{fig:age} lower left).
We do not have an explanation, but in the following paragraph
speculate on one possible line of future work that might help resolve the
discrepancy.

Analyzing Figure~\ref{fig:lithium}, it is remarkable that in just 50
million years, stars between $4500$\,K and $5200$\,K go from having
such a tight lithium sequence to one with a dispersion
$\approx10\times$ greater.  The existence of such a dispersion in
Pleiades-age K-dwarfs has been known for longer than some of the
present authors have been alive; it has also been known that the stars
with the most lithium are also the most rapidly rotating
\citep{butler_pleiades_1987,soderblom_evolution_1993}.  More recent
analyses of this correlation have been reviewed by
\citet{bouvier_lithium-rotation_2020}.  The conclusion of that work
was that the origin of the rotation-lithium correlation likely lies
within pre-main-sequence stellar physics.  If so, one would expect the
IC\,2062 and Tuc-Hor K-dwarfs to show a larger intrinsic lithium
dispersion, which they do not.  It seems therefore that the exact
onset time of the K-dwarf lithium dispersion might be worth
clarifying.  \citet{messina_rotation-lithium_2016} for instance do not
obviously see it in the $\beta$~Pic moving group ($25\pm3$\,Myr).
Given that RSG-5 seems somewhat isochronally older than the
$\delta$~Lyr cluster (and therefore IC\,2602
\citealt{bouma_kep1627_2022}), it might be a worthy cluster for such
an analysis.

\subsection{H$\alpha$}

As shown in Figure~\ref{fig:koiyouthindicators}, H$\alpha$ is in
emission for both components of KOI-7913, and in absorption for the
hotter stars.  Additionally, the emission appears double-peaked for
both of the KOI-7913 components.  We note that KOI-7913~A and
KOI-7913~B were spatially resolved from each other during data
acquisition.  Performing a cross-correlation between each of the stars
and the nearest matches in the Keck/HIRES template library, we also
found that the CCFs for both components of KOI-7913 showed no
indications of double-lined binarity \citep{kolbl_detection_2015}.

Balmer line emission, particularly for H$\alpha$, is expected for
stars of this age.  \citet{kraus_stellar_2014} for instance, in their
survey of Tuc-Hor ($\approx$40\,Myr), observed that essentially all
cluster members with spectral type $>$${\rm K4.5V}$ had H$\alpha$ in
emission.  The agrees with our observations that KOI-7913 shows
H$\alpha$ in emission for both components, while it is in absorption
for all of our other Kepler objects.  The double-peaked nature of the
emission, though not always present, is also not particularly rare for
active stars.  Proxima Centauri, for instance, has double-peaked
H$\alpha$ emission \citep{collins_calculations_2017}.  A simple
explanation is self-absorption: photons nearest the center of the line
see a greater optical depth from higher layers of the chromosphere,
while photons on the wings are on average too far from the
rest-wavelength to excite electrons and be re-absorbed in the upper
layers.  The exact details of when a star's atmosphere reaches the
conditions for such self-absorption require non-local thermal
equilibrium models of the chromosphere
\citep{short_chromospheric_1998,2005A&A...439.1137F}.

%Regarding indicators of youth, the primary has a spectral type
%$\approx$K6V, while the secondary is $\approx$K8V.  At $\approx$40
%Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
%but might be present in the primary \citep[{\it
%	e.g.},][Figure~8]{soderblom_ages_2014}.  Conversely, H$\alpha$ is
%expected to be in emission for for all M-dwarfs younger than
%$\approx150$\,Myr \citep[{\it e.g.},][]{kiman_calibration_2021}.  For
%the primary, H$\alpha$ emission is also expected based on
%previous spectroscopic studies of the Pleiades, IC~2602, and IC~2391
%\citep[{\it e.g.},][their Figure~8]{stauffer_rotational_1997}.  Based
%on the data from K.~Rocio et al's analysis\footnote{Available at
%	\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V
%star should be emitting in H$\alpha$ at 40\,Myr, but not typically
%past $\approx200\,{\rm Myr}$.  This presents a fairly direct test for
%the star's age.
%We

\section{Table of Transit Fit Parameters}
\label{app:transit}

Tables~\ref{tab:kepler1643}, \ref{tab:koi7368}, \ref{tab:koi7913} give
the full set of fitted and derived parameters from the model described in
Section~\ref{sec:fitting}.
Priors and convergence statistics are also listed.

\section{Disposition History of KOI-7913}
\label{app:koi7913}

The disposition of KOI-7913.01 has been debated: in
\texttt{q1\_q17\_dr25\_koi} the source was flagged as a false
positive, with the comment ``cent\_kic\_pos---halo\_ghost''.  This
comment and disposition were removed in the
\texttt{q1\_q17\_dr25\_sup\_koi} data release, which renamed the
planet a ``candidate''.  In this note, we discuss the interpretation
of these flags (which do not apply to the system, according to the
latest analysis).  We also discuss how the relative locations of
KOI-7913~A and KOI-7913~B affect the interpretation of the Kepler
data.

As described by \citet{thompson_planetary_2018}, the
``cent\_kic\_pos'' flag is an indication that the measured source
centroid is offset from its expected location in the Kepler Input
Catalog.  The final Kepler data validation reports, generated 2016 Jan
30, do not show this to be the case for KOI-7913.  Moreover, the
statistical significance of any centroid offset is lower than for
KOI-7368 and Kepler-1643 (which both show centroid offsets that are
likely explained by the stellar variability).

What of the ``halo\_ghost'' flag?  This test measures the transit
strength for the pixels inside the aperture, and compares it to that
measured in the ring of pixels around said aperture (the ``halo'').
Generally one expects the transit signal to be strongest in the
central aperture, rather than the halo.  Two types of false positive
scenarios can change this and trigger the flag: the first is when
optical ghosts from bright eclipsing binaries reflect off the CCD, and
contaminate the target star.  The second is when the PRF of nearby
stars directly overlaps with the PRF of the target star (see
\citealt{thompson_planetary_2018}, Section A.5.2).
The most obvious explanation for KOI-7913 is the latter case, given
that KOI-7913 B is $\approx0.9$ Kepler pixels away from Kepler-7913 A
and so it usually part of the ``halo''.  Due to the on-sky orientation
of KOI-7913 A and KOI-7913 B, the default ``optimal aperture''
selected in quarters 3, 7, 11, and 15 in fact included both stars,
while for the remaining quarters KOI-7913 B was excluded from the
optimal aperture but was included as part of the halo (see pages 35
through 71 of the data validation reports.)

Given the orientation of the stars and the $\approx$1.5 pixel FWHM of
the Kepler pixel response function, some blending between the two
stars is present.  The pointing geometries from quarters 3, 7, 11, and
15 however did not noticeably affect the observed transit depths,
which is an indication that the crowding metric applied in the data
products accurately correct the mean flux level
\citep{2017ksci.rept....6M}.  Analysis of the target-pixel data that
was acquired separately for KOI-7913~B also reveals a clearly
different stellar rotation period, and no hint of the transit signal.

\input{source_Kepler_1643_simpletransit_posteriortable.tex}
\input{source_KOI_7368_simpletransit_posteriortable.tex}
\input{source_KOI_7913_simpletransit_posteriortable.tex}

%\listofchanges
%\allauthors
\end{document}